### 2014-02-26 progress-report/related_work.tex
Added tex project for the progress report. Includes makefiles that build the pdf. The related works sections is a straight copy from Dominik's CSE552 report. Project plan currently empty.

\section{Related work}

Twitter developed Ambrose\cite{ambrose}, a platform for visualization and real-time monitoring of MapReduce data workflows. They offer three different views to show associated jobs, job dependencies and progress. Ambrose has been released as open source. Ambrose, however, is not suitable for our needs as the abstraction level of jobs is too high and does not capture single operators.

Google's Dapper\cite{sigelman2010dapper} is a distributed systems tracing infrastructure and offers fine grained tracing of calls in Google's distributed systems. They also proposed an interface for visualizing traces. Twitter closely modeled Zipkin\cite{zipkin} after Dapper and released it as open source. Our system is similar to Dapper and Zipkin as it has a similar architecture and offers similar visualizations. However, we offer different abstraction levels, which enables users to find problems faster and handle larger amounts of profiling data. Also, our visualization is specifically designed to help developers understand query execution in distributed database system like Myria as opposed to general traces in distributed systems.

Tools to visualize query plans for example in SQL Server, as used to improve performance for the SDSS Sky survey\cite{szalay2002sdss}, focus on optimizing queries and not query execution and have no visualization of data flow, which is necessary to optimize physical query execution.

### 2014-02-26 progress-report/report.tex
Added tex project for the progress report. Includes makefiles that build the pdf. The related works sections is a straight copy from Dominik's CSE552 report. Project plan currently empty.

\documentclass{article}
\usepackage{graphicx}
\usepackage{comment}
\usepackage{natbib}
\usepackage{listings}
\lstset{
    language=C,
    emph={nn_send, nn_receive, nn_configure, nn_invoke},
    emphstyle=\bfseries,
    basicstyle=\ttfamily\footnotesize,
    aboveskip=1mm plus 1mm minus 1mm,
    belowskip=1mm plus 1mm minus 1mm,
    mathescape=true,
    xleftmargin=\parindent,
}

\newcommand{\code}{\lstinline[basicstyle=\ttfamily\small]}

\newcommand{\todo}[2]{
\ifx\PrintTodo\undefined
\textcolor{red}{#2\xspace}
\else
\textcolor{red}{TODO[#1]:} \textcolor{blue}{\textit{#2}\xspace}
\fi
}

\begin{document}

\title{Visualizing Physical Query Execution in a Relational Big Data Management System}
\author{Umar Javed, Thierry Moreau, Dominik Moritz, Adriana Szekeres}

\maketitle

\input{related_work}
\input{project_plan}

\bibliographystyle{plain}
\bibliography{report}

\end{document}

### 2014-02-26 progress-report/abstract.tex
Added intro, started writing up the project planning paragraph.

\begin{abstract}

We propose a visualization system for understanding and exploring query execution and data movement in a distributed database system. Our tool will provide insight into: (1) data flow between query fragments and between workers, (2) query execution and operator dependencies, (3) cluster utilization, (4) network utilization. Our tool will help developers understand and improve query execution, and will provide insight into common problems such as data skew or performance bottlenecks.

In particular, our visualization system is built to inspect query execution in Myria, a distributed big data management system currently being developed in the UW CSE database group. Myria aims towards building a distributed database platform to provide \emph{big data management and analytics as a service} primarily for scientific applications.

Our vision is to build a generic visualization that can be used for other systems as well (e.g. Spark, Hadoop). This project builds on an existing prototype that uses logs generated during query execution to create a time series visualization of operator states.

\end{abstract}

### 2014-02-26 progress-report/project_plan.tex
Added intro, started writing up the project planning paragraph.

\section{Project Plan}

Our team of four is led by Dominik, who started working on a query visualization framework for Myria last quarter. The responsibilities for each of the group members are listed below:

\begin {itemize}
    \item Umar Javed: visualizing network data flow between workers
    \item Thierry Moreau: visualizing query plan
    \item Dominik Moritz: team leader, in charge of data collection and back end
    \item Adriana Szerkes: visualizing query fragment execution
\end{itemize}

### 2014-02-26 progress-report/project_plan.tex
Added project milestones
Add paragraph about current prototype.

@@
 \section{Project Plan}
+
+\subsection{Team Responsibilities}
 
@@
     \item Thierry Moreau: visualizing query plan
-    \item Dominik Moritz: team leader, in charge of data collection and back end
+    \item Dominik Moritz: architect, data collection and back end
     \item Adriana Szerkes: visualizing query fragment execution
 \end{itemize}
+
+\subsection{Milestones}
+
+\paragraph{Current Prototype}
+
+Dominik built a prototype last quarter for the \emph{Distributed Systems} class.
+The goal of this project was to build a system to collect and analyze logs in the Myria \footnote{\url{http://myria.cs.washington.edu/}} system.
+The prototype allows a user to switch between three levels of query execution visualization to identify the problem at the granularity that best reveals the information they need.
+The goal for this project is to improve the visualization part of the previous prototype and build a more consistent exploration tool.
+Furthermore, the back end has to be rewritten to support visualization of long running queries in the Myria system but are not supported in the current prototype.
+
+\paragraph{Milestones}
+
+\begin{enumerate}
+    \item Project Planning (completed Feb 23rd)
+    \begin{enumerate}
+        \item Brainstorming and goals
+        \item List questions we'd like to answer
+        \item Develop a consistent story-board
+        \item Assign responsibilities
+    \end{enumerate}
+    \item Initial System Setup \& Specifications (completed Feb 23rd)
+    \begin{enumerate}
+        \item Set up cluster \& web front-end
+        \item Data collection from query-runs
+        \item Formalize interfaces and data formats
+    \end{enumerate}
+
+\end{enumerate}

### 2014-02-26 progress-report/project_plan.tex
added more milestones

--- 2014-02-26 progress-report/project_plan.tex
@@
     \item Thierry Moreau: visualizing query plan
-    \item Dominik Moritz: team leader, in charge of data collection and back end
+    \item Dominik Moritz: architect, data collection and back end
     \item Adriana Szerkes: visualizing query fragment execution
 \end{itemize}
+
+\subsection{Milestones}
+
+\paragraph{Current Prototype}
+
+\paragraph{Milestones}
+
+\begin{enumerate}
+    \item Project Planning (completed Feb 23rd)
+    \begin{enumerate}
+        \item Brainstorming and goals
+        \item List questions we'd like to answer
+        \item Develop a consistent story-board
+        \item Assign responsibilities
+    \end{enumerate}
+    \item Initial System Setup \& Specifications (completed Feb 23rd)
+    \begin{enumerate}
+        \item Set up cluster \& web front-end
+        \item Data collection from query-runs
+        \item Formalize interfaces and data formats
+    \end{enumerate}
+    \item Initial Prototype (complete by March 5th)
+    \begin{enumerate}
+        \item Produce initial visualizations
+        \item Add a basic interaction between views
+    \end{enumerate}
+    \item Working Prototype (complete by March 13th)
+    \begin{enumerate}
+        \item Improve visualizations and interaction
+    \end{enumerate}
+    \item Project Deliverables 
+    \begin{enumerate}
+        \item Produce poster (complete by March 13th)
+        \item Complete demo (complete by March 13th)
+        \item Write the project report (complete by March 20th)
+    \end{enumerate}
+
+\end{enumerate}

### 2014-02-27 progress-report/abstract.tex
move footnote

@@
 
-In particular, our visualization system is built to inspect query execution in Myria, a distributed big data management system currently being developed in the UW CSE database group.
+In particular, our visualization system is built to inspect query execution in Myria\footnote{\url{http://myria.cs.washington.edu/}}, a distributed big data management system currently being developed in the UW CSE database group.
 Myria aims towards building a distributed database platform to provide \emph{big data management and analytics as a service} primarily for scientific applications.

### 2014-02-27 progress-report/project_plan.tex
move footnote

--- 2014-02-26 progress-report/project_plan.tex
@@
 Dominik built a prototype last quarter for the \emph{Distributed Systems} class.
-The goal of this project was to build a system to collect and analyze logs in the Myria \footnote{\url{http://myria.cs.washington.edu/}} system.
+The goal of this project was to build a system to collect and analyze logs in the Myria system.
 The prototype allows a user to switch between three levels of query execution visualization to identify the problem at the granularity that best reveals the information they need.
@@
     \end{enumerate}
-
+    \item Initial Prototype (complete by March 5th)
+    \begin{enumerate}
+        \item Produce initial visualizations
+        \item Add a basic interaction between views
     \end{enumerate}
+    \item Working Prototype (complete by March 13th)
+    \begin{enumerate}
+        \item Improve visualizations and interaction
+    \end{enumerate}
+    \item Project Deliverables 
+    \begin{enumerate}
+        \item Produce poster (complete by March 13th)
+        \item Complete demo (complete by March 13th)
+        \item Write the project report (complete by March 20th)
+    \end{enumerate}
+\end{enumerate}

### 2014-03-16 final-report/makepaper.sh
Initial template for final report

@@ -1,0 +1,5 @@
+#! /bin/bash
+pdflatex paper
+bibtex paper
+pdflatex paper
+pdflatex paper

### 2014-03-16 final-report/paper.tex
Started outline

\documentclass{chi2009}
\usepackage{times}
\usepackage{url}
\usepackage{graphics}
\usepackage{color}
\usepackage[pdftex]{hyperref}
\hypersetup{%
pdftitle={Your Title},
pdfauthor={Your Authors},
pdfkeywords={your keywords},
bookmarksnumbered,
pdfstartview={FitH},
colorlinks,
citecolor=black,
filecolor=black,
linkcolor=black,
urlcolor=black,
breaklinks=true,
}
\newcommand{\comment}[1]{}
\definecolor{Orange}{rgb}{1,0.5,0}
\newcommand{\todo}[1]{\textsf{\textbf{\textcolor{Orange}{[[#1]]}}}}

\pagenumbering{arabic}  % Arabic page numbers for submission.  Remove this line to eliminate page numbers for the camera ready copy

\begin{document}
% to make various LaTeX processors do the right thing with page size
\special{papersize=8.5in,11in}
\setlength{\paperheight}{11in}
\setlength{\paperwidth}{8.5in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

% use this command to override the default ACM copyright statement 
% (e.g. for preprints). Remove for camera ready copy.
%\toappear{Submitted for review to CHI 2009.}
\toappear{}

\title{The Title}
\numberofauthors{2}
\author{
  \alignauthor Author 1\\
    \affaddr{Affiliation}\\
    \affaddr{Affiliation}\\
    \email{author@a.com}
  \alignauthor Author 2\\
    \affaddr{Affiliation}\\
    \affaddr{Affiliation}\\
    \email{author2@b.com}
}

\maketitle

\begin{abstract}
  Abstract.
\end{abstract}

%\keywords{put author keywords here}

%\category{H.5.2}{Information Interfaces and Presentation}{Miscellaneous}[Optional sub-category]

\section{Introduction}

\begin{itemize}
    \item problem: hard to learn why a query is slow
    \item motivation: help developers/programmers dig what the problems might be
\end{itemize}

\section{Related Work}

\section{Approach}

\begin{itemize}
    \item determine possible causes (data skew, stragglers, etc.) for why a query might be slow and design visualizations to
help 
    \item we designed three views - physical query plan, fragment view, communication view.
\end{itemize}

\subsection{Physical query plan view}
\subsection{Fragment execution view}
\subsection{Worker communication view}

\section{Evaluation}

\section{Conclusion}

\section{Future Work}

\bibliographystyle{abbrv}
\bibliography{sample}

\end{document}

### 2014-03-16 final-report/paper.tex
Assigned people
Abstract and authors

@@
 \hypersetup{%
-pdftitle={Your Title},
-pdfauthor={Your Authors},
-pdfkeywords={your keywords},
+pdftitle={Visualizing Physical Query Execution in a Relational Big Data Management System},
+pdfauthor={Umar Javed, Thierry Moreau, Dominik Moritz, Adriana Szekeres},
+pdfkeywords={},
 bookmarksnumbered,
@@
 
-\title{The Title}
-\numberofauthors{2}
-\author{
-  \alignauthor Author 1\\
-    \affaddr{Affiliation}\\
-    \affaddr{Affiliation}\\
-    \email{author@a.com}
-  \alignauthor Author 2\\
-    \affaddr{Affiliation}\\
-    \affaddr{Affiliation}\\
-    \email{author2@b.com}
+\title{Visualizing Physical Query Execution in a Relational Big Data Management System}
+\numberofauthors{1}
+\author{\alignauthor Umar Javed, Thierry Moreau, Dominik Moritz, Adriana Szekeres \\
+\affaddr{Dept. of Computer Science, University of Washington} \\ \affaddr{ Seattle, Washington, USA} \\
+\email{\texttt{\{ujaved, moreau, domoritz, aaasz\}@cs.washington.edu}}
 }
@@
 \begin{abstract}
-  Abstract.
+We propose a visualization system for understanding and exploring query execution and data movement in a distributed database system.
+Our tool will provide insight into: (1) data flow between query fragments and between workers, (2) query execution and operator dependencies, (3) cluster utilization, (4) network utilization.
+Our tool will help developers understand and improve query execution, and will provide insight into common problems such as data skew or performance bottlenecks.
+
+In particular, our visualization system is built to inspect query execution in Myria\footnote{\url{http://myria.cs.washington.edu/}}, a distributed big data management system currently being developed in the UW CSE database group.
+Myria aims towards building a distributed database platform to provide \emph{big data management and analytics as a service} primarily for scientific applications.
+
+Our vision is to build a generic visualization that can be used for other systems as well (e.g.
+Spark, Hadoop).
+This project builds on an existing prototype that uses logs generated during query execution to create a time series visualization of operator states.
+
 \end{abstract}
@@
 
+% Umar
+
 \section{Related Work}
 
+% Dom
+
 \section{Approach}
+
+% Adriana
+
+% What the user sees
+% Why did we choose the visualization
+% How it is implemented
 
@@
 
-\subsection{Physical query plan view}
-\subsection{Fragment execution view}
-\subsection{Worker communication view}
+\subsection{Back-end}
+
+% Dom
+
+\subsection{Front-end}
+
+% Thierry
+
+\subsubsection{Physical query plan view}
+
+% Thierry
+
+\subsubsection{Fragment execution view}
+
+% Adriana
+
+\subsubsection{Worker communication view}
+
+% Umar
 
@@
 
+% How it helps
+% Examples
+
+% Thierry + Dom
+
+\textbf{Discussion:}
+
 \section{Conclusion}
 
+% Adriana
+
 \section{Future Work}
+
+% Dom
 

### 2014-03-16 final-report/paper.tex
Related work

@@
 \begin{abstract}
+
 We propose a visualization system for understanding and exploring query execution and data movement in a distributed database system.
@@
 
+% What is myria, how does it work (look at report from last quarter but shorter)
+% motivation
+% questions
+% why is a visualization a good way to answer the questions
+
 \begin{itemize}
@@
 
-% Dom
+Twitter developed Ambrose\cite{ambrose}, a platform for visualization and real-time monitoring of MapReduce data workflows.
+They offer three different views to show associated jobs, job dependencies and progress.
+Ambrose has been released as open source.
+Ambrose, however, is not suitable for our needs as the abstraction level of jobs is too high and does not capture single operators.
+
+Google's Dapper\cite{sigelman2010dapper} is a distributed systems tracing infrastructure and offers fine grained tracing of calls in Google's distributed systems.
+They also proposed an interface for visualizing traces.
+Twitter closely modeled Zipkin\cite{zipkin} after Dapper and released it as open source.
+Our system is similar to Dapper and Zipkin as it has a similar architecture and offers similar visualizations.
+However, we offer different abstraction levels, which enables users to find problems faster and handle larger amounts of profiling data.
+Also, our visualization is specifically designed to help developers understand query execution in distributed database system like Myria as opposed to general traces in distributed systems.
+
+Tools to visualize query plans for example in SQL Server, as used to improve performance for the SDSS Sky survey\cite{szalay2002sdss}, focus on optimizing queries and not query execution and have no visualization of data flow, which is necessary to optimize physical query execution.
 
@@
 % Thierry
+
+\subsubsection{Overview over all fragments}
+
+% Adriana
+
+% Write about small multiples view
 

### 2014-03-16 final-report/paper.tex
added paper macro
x trace related work

@@
 \newcommand{\todo}[1]{\textsf{\textbf{\textcolor{Orange}{[[#1]]}}}}
+\newcommand{\system}{OGEN }
 
@@
 They also proposed an interface for visualizing traces.
-Twitter closely modeled Zipkin\cite{zipkin} after Dapper and released it as open source.
-Our system is similar to Dapper and Zipkin as it has a similar architecture and offers similar visualizations.
-However, we offer different abstraction levels, which enables users to find problems faster and handle larger amounts of profiling data.
-Also, our visualization is specifically designed to help developers understand query execution in distributed database system like Myria as opposed to general traces in distributed systems.
+Similarly, X-trace\cite{fonseca2007x} was developed as a framework to trace which events cause what other events in a distributed environment.
+recently, there has been work on visualizing event traces collected in X-trace\footnote{\url{https://github.com/brownsys/X-Trace/tree/master/src/webui/html/interactive}}.
+Due to the importance of these kinds of debug facilities, Twitter closely modeled Zipkin\cite{zipkin} after Dapper and X-trace and released it as open source.
+
+Dapper and X-trace focus on how data flows through a distributed system.
+In \system the user looks at the operators and how the data flows through them.
+This orthogonal view is better suited for debugging performance bottlenecks in distributed DBMS and also scales to a larger number of events.
+Furthermore, in contrast to  we offer different abstraction levels, which enables users to find problems faster and handle larger amounts of profiling data.
+Also, \system is specifically designed to help developers understand query execution in distributed database system like Myria as opposed to general traces in distributed systems.
 
@@
 
-% Thierry
-
 \subsubsection{Overview over all fragments}
@@
 % Thierry
+
+\system is awesome.
 
@@
 \bibliographystyle{abbrv}
-\bibliography{sample}
+\bibliography{paper}
 

### 2014-03-16 final-report/paper.tex
Xspace

@@
 \usepackage[pdftex]{hyperref}
+\usepackage{xspace}
+
+\newcommand*{\system}{OGEN\xspace}
+\newcommand*{\papertitle}{\system: Visualizing Physical Query Execution in a Relational Big Data Management System}
+
 \hypersetup{%
-pdftitle={Visualizing Physical Query Execution in a Relational Big Data Management System},
+pdftitle={\papertitle},
 pdfauthor={Umar Javed, Thierry Moreau, Dominik Moritz, Adriana Szekeres},
@@
 \newcommand{\todo}[1]{\textsf{\textbf{\textcolor{Orange}{[[#1]]}}}}
-\newcommand{\system}{OGEN }
 
@@
 
-\title{Visualizing Physical Query Execution in a Relational Big Data Management System}
+\title{\papertitle}
 \numberofauthors{1}
@@
 Dapper and X-trace focus on how data flows through a distributed system.
-In \system the user looks at the operators and how the data flows through them.
+In \system the visualization focuses on the operators and how the data flows through them.
 This orthogonal view is better suited for debugging performance bottlenecks in distributed DBMS and also scales to a larger number of events.

### 2014-03-16 final-report/paper.tex
Better

@@
 
-We propose a visualization system for understanding and exploring query execution and data movement in a distributed database system.
+We propose a visualization system for understanding and exploring query execution and data movement in a distributed database management system (DDBMS).
 Our tool will provide insight into: (1) data flow between query fragments and between workers, (2) query execution and operator dependencies, (3) cluster utilization, (4) network utilization.
-Our tool will help developers understand and improve query execution, and will provide insight into common problems such as data skew or performance bottlenecks.
+Our tool, which we call \system, will help developers understand and improve query execution, and will provide insight into common problems such as data skew or performance bottlenecks.
 
-In particular, our visualization system is built to inspect query execution in Myria\footnote{\url{http://myria.cs.washington.edu/}}, a distributed big data management system currently being developed in the UW CSE database group.
+In particular, \system is built to inspect query execution in Myria\footnote{\url{http://myria.cs.washington.edu/}}, a distributed big data management system currently being developed in the UW CSE database group.
 Myria aims towards building a distributed database platform to provide \emph{big data management and analytics as a service} primarily for scientific applications.
 
-Our vision is to build a generic visualization that can be used for other systems as well (e.g.
-Spark, Hadoop).
-This project builds on an existing prototype that uses logs generated during query execution to create a time series visualization of operator states.
+The proposed visualizations can easily be applied to other DDBMS as well (e.g. Spark, Hadoop).
 
@@
 \section{Introduction}
+
+% Umar
 
@@
 \end{itemize}
-
-% Umar
 
@@
 In \system the visualization focuses on the operators and how the data flows through them.
-This orthogonal view is better suited for debugging performance bottlenecks in distributed DBMS and also scales to a larger number of events.
+This orthogonal view is better suited for debugging performance bottlenecks in DDBMS and also scales to a larger number of events.
 Furthermore, in contrast to  we offer different abstraction levels, which enables users to find problems faster and handle larger amounts of profiling data.
@@
 
+% Thierry
+
 \subsubsection{Overview over all fragments}
@@
 % Thierry
-
-\system is awesome.
 

### 2014-03-16 final-report/paper.tex
added to Approach intro

--- 2014-03-16 final-report/paper.tex
@@
 
-% What is myria, how does it work (look at report from last quarter but shorter)
+% What is myria, how does it work (look at report from last quarter but shorter), which languages are supported
 % motivation
@@
 They also proposed an interface for visualizing traces.
-Similarly, X-trace\cite{fonseca2007x} was developed as a framework to trace which events cause what other events in a distributed environment.
-recently, there has been work on visualizing event traces collected in X-trace\footnote{\url{https://github.com/brownsys/X-Trace/tree/master/src/webui/html/interactive}}.
-Due to the importance of these kinds of debug facilities, Twitter closely modeled Zipkin\cite{zipkin} after Dapper and X-trace and released it as open source.
-
-Dapper and X-trace focus on how data flows through a distributed system.
-In \system the user looks at the operators and how the data flows through them.
-This orthogonal view is better suited for debugging performance bottlenecks in distributed DBMS and also scales to a larger number of events.
-Furthermore, in contrast to  we offer different abstraction levels, which enables users to find problems faster and handle larger amounts of profiling data.
-Also, \system is specifically designed to help developers understand query execution in distributed database system like Myria as opposed to general traces in distributed systems.
+Twitter closely modeled Zipkin\cite{zipkin} after Dapper and released it as open source.
+Our system is similar to Dapper and Zipkin as it has a similar architecture and offers similar visualizations.
+However, we offer different abstraction levels, which enables users to find problems faster and handle larger amounts of profiling data.
+Also, our visualization is specifically designed to help developers understand query execution in distributed database system like Myria as opposed to general traces in distributed systems.
 
@@
 
+\system's frontend is embedded into Myria's web frontend and provides comprehensive performance views to the user.
+There are 4 views in total that allow the user to visualize query performance details under different angles:
+\begin{itemize}
+    \item Physical Query Plan view: this view allows the user to connect the query she wrote with the physical query plan that the query optimizer generates.
+    \item etc. 
+\end{itemize}
+
 \subsection{Back-end}
@@
 \bibliographystyle{abbrv}
-\bibliography{paper}
+\bibliography{sample}
 

### 2014-03-16 final-report/paper.tex

--- 2014-03-16 final-report/paper.tex
@@
     \item we designed three views - physical query plan, fragment view, communication view.
+\end{itemize}
+
+\system's frontend is embedded into Myria's web frontend and provides comprehensive performance views to the user.
+There are 4 views in total that allow the user to visualize query performance details under different angles:
+\begin{itemize}
+    \item Physical Query Plan view: this view allows the user to connect the query she wrote with the physical query plan that the query optimizer generates.
+    \item etc. 
 \end{itemize}

### 2014-03-16 final-report/paper.tex
Added macros

@@
 \newcommand*{\papertitle}{\system: Visualizing Physical Query Execution in a Relational Big Data Management System}
+\newcommand*{\graphview}{Physical Query Plan View\xspace}
+\newcommand*{\fragmentview}{Fragment Execution View\xspace}
+\newcommand*{\networkview}{Worker Communication View\xspace}
+\newcommand*{\overallview}{Fragment Overall View\xspace}
 
@@
 
-% What the user sees
+The first step in designing \system was to identify the possible causes that
+could affect the execution performance of the query. After we identified
+several such causes, that we will enumerate below, we leveraged visualization
+techniques that we found fit to allow developers/programmers to get insight
+into how the query was executed on the available resources and quickly
+identify the source(s) of the performance penalty.
+
+We found that the performance of the query might be affected by the following
+factors, which influenced :
+\begin{itemize}
+   \item 
+\end{itemize}
+
 % Why did we choose the visualization
-% How it is implemented
 

### 2014-03-16 final-report/paper.tex
front-end WIP

--- 2014-03-16 final-report/paper.tex
@@
 \subsection{Front-end}
-
+The Myria web front-end server is written in Python and runs on Google App Engine.
+\system's user-interface (UI) is embedded into Myria's web front-end.
+We build \system's UI using D3 to delive an interactive visualization experience to the user.
+D3 is a JavaScript framework for data visualziation on the web.
+We used visualization techniques such as empth{focus+context} that are possible to implement conveniently using D3.
+The \system's front-end gets data from the Myria server in JSON and CSV file formats.
+JSON objects are used to describe the physical queryplan, while CSV files contain information describing the execution of the query at the fragment and operator level, the total amount of data exchange between workers at different stages of the query execution.
+The \system's web UI is divided into two main sections.
 % Thierry

### 2014-03-17 final-report/paper.tex
macros
Fixes
Describe backend

@@
 
+\usepackage{graphicx}
+
 \newcommand*{\system}{OGEN\xspace}
 \newcommand*{\papertitle}{\system: Visualizing Physical Query Execution in a Relational Big Data Management System}
-\newcommand*{\graphview}{Physical Query Plan View\xspace}
-\newcommand*{\fragmentview}{Fragment Execution View\xspace}
-\newcommand*{\networkview}{Worker Communication View\xspace}
-\newcommand*{\overallview}{Fragment Overall View\xspace}
+\newcommand*{\graph}{Physical Query Plan\xspace}
+\newcommand*{\fragment}{Fragment Execution\xspace}
+\newcommand*{\network}{Worker Communication\xspace}
+\newcommand*{\overall}{Fragment Overall\xspace}
 
@@
 % Why did we choose the visualization
+% How it is implemented
+% explain how events and logs are used to create the raw data for vis
 
@@
 
+The front and back end of \system are highly decoupled so that it can be used not only with Myria but also with other systems such as Spark or Hadoop.
+In Section~\ref{sec:back} we illustrate how we implemented log collections and aggregation in Myria.
+
 \subsection{Back-end}
+\label{sec:back}
 
-% Dom
+There are two possible approaches that we explored to collect event logs that can be used to extract spans and other performance related data necessary for the visualization.
+First, events could be logged to files using a standard logging tool and then collect the files on one node using remote copy.
+Once the logs are in one place, they can be parsed and the data transformed into the desired format.
+
+We switched to a different approach as shown in Figure~\ref{fig:arch} where logs are written directly into the worker's database.
+When the front-end requests the data, the master executes a query that collects the data and streams the results directly to the requesting client.
+Logs are thus tuples in a relation and no longer separate in files and in a custom format.
+Despite being more reliable, faster and using the same logic that Myria already implements, we are able to write queries to aggregate, filter and transform the collected data.
+
+For example, the data for how many workers are working on a certain fragment as described in Section~\ref{sec:fragments} is generated from the even log using a query to select only events for the root operators (the operators that have no parent in the same fragment) in each fragment and a custom map function that carries state.
+The state is an integer that is incremented when the operator is called on a worker and decremented when the operator returns.
+
+\begin{figure}[ht]
+  \includegraphics[width=\columnwidth]{viz_arch}
+  \caption{Architecture}
+  \label{fig:arch}
+\end{figure}
+
+We highly suggest that developers use the second approach when porting \system to a new DDBMS.
 
@@
 \subsubsection{Overview over all fragments}
+\label{sec:fragments}
 
@@
 % Adriana
+
+% use the word spans when referring to events with duration (used in dapper)
 

### 2014-03-17 final-report/paper.tex
More to approach

--- 2014-03-16 final-report/paper.tex
@@
 \newcommand*{\papertitle}{\system: Visualizing Physical Query Execution in a Relational Big Data Management System}
-\newcommand*{\graphview}{Physical Query Plan View\xspace}
-\newcommand*{\fragmentview}{Fragment Execution View\xspace}
-\newcommand*{\networkview}{Worker Communication View\xspace}
-\newcommand*{\overallview}{Fragment Overall View\xspace}
+\newcommand*{\graph}{Physical Query Plan\xspace}
+\newcommand*{\fragment}{Fragment Execution\xspace}
+\newcommand*{\network}{Worker Communication\xspace}
+\newcommand*{\overall}{Fragment Overall\xspace}
 
@@
 We found that the performance of the query might be affected by the following
-factors, which influenced :
+factors, which greatly influenced the design of our visualization tool:
 \begin{itemize}
-   \item 
+   \item \emph{Wrong/unoptimized physical query plan.} We manually analyzed
+several physical query plans and discovered that some queries were poorly
+optimized. This led us to design a view that allows interactive exploration of
+the physical query plan (mapped to a graph of query operators). 
+   \item Stragglers, tail latency, expensive query operators, poor scheduling.
+   \item Poor data partitioning, data skew.
 \end{itemize}

### 2014-03-17 final-report/paper.tex
front end section

--- 2014-03-17 final-report/paper.tex
@@
 We found that the performance of the query might be affected by the following
-factors, which greatly influenced the design of our visualization tool:
+factors, which influenced :
 \begin{itemize}
-   \item \emph{Wrong/unoptimized physical query plan.} We manually analyzed
-several physical query plans and discovered that some queries were poorly
-optimized. This led us to design a view that allows interactive exploration of
-the physical query plan (mapped to a graph of query operators). 
-   \item Stragglers, tail latency, expensive query operators, poor scheduling.
-   \item Poor data partitioning, data skew.
+   \item 
 \end{itemize}
@@
 D3 is a JavaScript framework for data visualziation on the web.
-We used visualization techniques such as empth{focus+context} that are possible to implement conveniently using D3.
-The \system's front-end gets data from the Myria server in JSON and CSV file formats.
-JSON objects are used to describe the physical queryplan, while CSV files contain information describing the execution of the query at the fragment and operator level, the total amount of data exchange between workers at different stages of the query execution.
-The \system's web UI is divided into two main sections.
+We used visualization techniques such as \emph{focus+context} that possible to implement conveniently using D3.
+
+\system's front-end gets data from the Myria server as JSON and CSV files.
+JSON objects are used to describe the physical queryplan.
+CSV files contain logs on query execution at the fragment and operator level as well as data exchange between workers at different stages of the query execution.
+
+\system's web UI is divided into two components: (1) the context window which provides a visualization of the \graph, and (2) the performance window which provides some insight into an element selected in the context view.
+The context window contains a view of the \graph, rendered as a graph.
+The user can navigate the \graph by expanding query fragments into the operators that compose it.
+The user can chose to select a fragment of interest which will render a \fragment visualization for the selected fragment in the performance window.
+Alternatively, the user can select an fragment-to-fragment edge in the \graph thus rendering a \network visualization for the selected edge in the performance window.
+If no elements are selected in the \graph view, an \overall visualization is rendered in the performance window, displaying the aggregate worker utilization over time for each fragment.
+
+The following subsections describe each one of the four views offered by \system's web UI.
+
 % Thierry

### 2014-03-17 final-report/paper.tex

@@
 \usepackage{xspace}
+
+\usepackage{graphicx}
 

### 2014-03-17 final-report/paper.tex
Intro for arch

@@
 \section{Approach}
+\label{sec:approach}
 
@@
 
+The task of the back-end is to log when operators are called, when the call returns and when data is sent to another worker.
+Using this data and simple transformations, we can obtain the information necessary to create the visualizations to answer the questions in Section~\ref{sec:approach}.
+
 There are two possible approaches that we explored to collect event logs that can be used to extract spans and other performance related data necessary for the visualization.
@@
 
-For example, the data for how many workers are working on a certain fragment as described in Section~\ref{sec:fragments} is generated from the even log using a query to select only events for the root operators (the operators that have no parent in the same fragment) in each fragment and a custom map function that carries state.
+For example, the data for how many workers are working on a certain fragment as described in Section~\ref{sec:fragments} is generated from the event log using a query to select only events for the root operators (the operators that have no parent in the same fragment) in each fragment and a custom map function that carries state.
 The state is an integer that is incremented when the operator is called on a worker and decremented when the operator returns.

### 2014-03-17 final-report/paper.tex
Fixes

--- 2014-03-17 final-report/paper.tex
@@
 We found that the performance of the query might be affected by the following
-factors, which influenced :
+factors, which greatly influenced the design of our visualization tool:
 \begin{itemize}
-   \item 
+   \item \emph{Wrong/unoptimized physical query plan.} We manually analyzed
+several physical query plans and discovered that some queries were poorly
+optimized. This led us to design a view that allows interactive exploration of
+the physical query plan (mapped to a graph of query operators). 
+   \item \emph{Stragglers, tail latency, expensive query operators, poor
+scheduling.} Our targeted systems are \emph{distributed} databases. Therefore,
+as the query is executed in a distributed environment, the query coordinator
+has to wait until all the workers finish executing their assigned tasks. To
+give insight into how the tasks are distributed and executed at each workers,
+we designed a view that shows the cluster utilization over time and at each
+worker.
+   \item \emph{Poor data partitioning, data skew.} Network traffic can cause
+serious delay when executing a distributed query. Due to poor data
+partitioning, workers might need to send unnecessary big amount of data between
+each other. We designed a view that allows the developer/programmer to analyze
+the data traffic generated while executing the query.  
 \end{itemize}
@@
 
-\begin{itemize}
-    \item determine possible causes (data skew, stragglers, etc.) for why a query might be slow and design visualizations to
-help
-    \item we designed three views - physical query plan, fragment view, communication view.
-\end{itemize}
-
-\system's frontend is embedded into Myria's web frontend and provides comprehensive performance views to the user.
-There are 4 views in total that allow the user to visualize query performance details under different angles:
-\begin{itemize}
-    \item Physical Query Plan view: this view allows the user to connect the query she wrote with the physical query plan that the query optimizer generates.
-    \item etc. 
-\end{itemize}
-
-The front and back end of \system are highly decoupled so that it can be used not only with Myria but also with other systems such as Spark or Hadoop.
-In Section~\ref{sec:back} we illustrate how we implemented log collections and aggregation in Myria.
+\system's architecture (Fig.~\ref{fig:arch}) consists of a back-end,
+used as a plug-in interface for the distributed database system targeted, and a
+front-end which implements the views described above. The front-end and
+back-end of \system are highly decoupled so that it can be used not only with
+Myria but also with other systems such as Spark or Hadoop. In
+Section~\ref{sec:back} we illustrate how we implemented log collections and
+aggregation in Myria.
 
@@
 D3 is a JavaScript framework for data visualziation on the web.
-We used visualization techniques such as \emph{focus+context} that possible to implement conveniently using D3.
-
-\system's front-end gets data from the Myria server as JSON and CSV files.
-JSON objects are used to describe the physical queryplan.
-CSV files contain logs on query execution at the fragment and operator level as well as data exchange between workers at different stages of the query execution.
-
-\system's web UI is divided into two components: (1) the context window which provides a visualization of the \graph, and (2) the performance window which provides some insight into an element selected in the context view.
-The context window contains a view of the \graph, rendered as a graph.
-The user can navigate the \graph by expanding query fragments into the operators that compose it.
-The user can chose to select a fragment of interest which will render a \fragment visualization for the selected fragment in the performance window.
-Alternatively, the user can select an fragment-to-fragment edge in the \graph thus rendering a \network visualization for the selected edge in the performance window.
-If no elements are selected in the \graph view, an \overall visualization is rendered in the performance window, displaying the aggregate worker utilization over time for each fragment.
-
-The following subsections describe each one of the four views offered by \system's web UI.
-
+We used visualization techniques such as empth{focus+context} that are possible to implement conveniently using D3.
+The \system's front-end gets data from the Myria server in JSON and CSV file formats.
+JSON objects are used to describe the physical queryplan, while CSV files contain information describing the execution of the query at the fragment and operator level, the total amount of data exchange between workers at different stages of the query execution.
+The \system's web UI is divided into two main sections.
 % Thierry

### 2014-03-17 final-report/paper.tex
Little fixes, added outlook

--- 2014-03-17 final-report/paper.tex
@@
 We propose a visualization system for understanding and exploring query execution and data movement in a distributed database management system (DDBMS).
-Our tool will provide insight into: (1) data flow between query fragments and between workers, (2) query execution and operator dependencies, (3) cluster utilization, (4) network utilization.
-Our tool, which we call \system, will help developers understand and improve query execution, and will provide insight into common problems such as data skew or performance bottlenecks.
+Our tool provides insight into: (1) data flow between query fragments and between workers, (2) query execution and operator dependencies, (3) cluster utilization, (4) network utilization.
+Our tool, which we call \system, helps developers understand and improve query execution, and provides insight into common problems such as data skew or performance bottlenecks.
 
@@
 First, events could be logged to files using a standard logging tool and then collect the files on one node using remote copy.
-Once the logs are in one place, they can be parsed and the data transformed into the desired format.
+Once the logs are on one machine, the logs can be parsed and the data can be transformed into the desired format.
 
@@
   \includegraphics[width=\columnwidth]{viz_arch}
-  \caption{Architecture}
+  \caption{Overview of the log collection and transformation architecture.
+Raw event logs are collected on each worker.
+The data is stored in relations.
+To download the data, a query has to be executed.}
   \label{fig:arch}
@@
 \subsection{Front-end}
+
 The Myria web front-end server is written in Python and runs on Google App Engine.
@@
 
-% Dom
+A focus of the current implementation is on scalability to many workers and long running queries.
+More aggregations have to be done by the database itself.
+Most importantly, in the fragment visualization that shows spans for all events we should limit the data that is being downloaded and rendered to what the user can see.
+With the architecture as explained in Section~\ref{sec:back}, we are able to express these filter conditions as queries.
 

### 2014-03-17 final-report/paper.tex
Added images folder
More future work

@@
 
+Besides performance improvements, we plan to integrate X-trace and its visualization to offer the orthogonal view and allow users to trace how tuple batches flow through the operators and between operators.
+Combined with the visualizations described in this paper, it will form a powerful debugging tool that will help us to better understand and improve query execution.
+
 \bibliographystyle{abbrv}

### 2014-03-17 final-report/paper.tex

--- 2014-03-17 final-report/paper.tex
@@
 \begin{figure}[ht]
-  \includegraphics[width=\columnwidth]{viz_arch}
-  \caption{Architecture}
+  \includegraphics[width=\columnwidth]{images/viz_arch}
+  \caption{Overview of the log collection and transformation architecture.
+Raw event logs are collected on each worker.
+The data is stored in relations.
+To download the data, a query has to be executed.}
   \label{fig:arch}
@@
 
+The \fragment view comprises two charts: the aggregation chart and the workers chart.
+
 % use the word spans when referring to events with duration (used in dapper)
@@
 
-% Dom
+A focus of the current implementation is on scalability to many workers and long running queries.
+More aggregations have to be done by the database itself.
+Most importantly, in the fragment visualization that shows spans for all events we should limit the data that is being downloaded and rendered to what the user can see.
+With the architecture as explained in Section~\ref{sec:back}, we are able to express these filter conditions as queries.
+
+Besides performance improvements, we plan to integrate X-trace and its visualization to offer the orthogonal view and allow users to trace how tuple batches flow through the operators and between operators.
+Combined with the visualizations described in this paper, it will form a powerful debugging tool that will help us to better understand and improve query execution.
 

### 2014-03-17 final-report/paper.tex
Small things and added d3 citation

@@
 % motivation
-% questions
+% questions (inspiration: https://docs.google.com/a/uw.edu/document/d/1Syn65bu_p6S4RQETaozsleYcRBauQJ7IduYrctS1wrw/edit?pli=1#)
 % why is a visualization a good way to answer the questions
@@
 front-end which implements the views described above. The front-end and
-back-end of \system are highly decoupled so that it can be used not only with
+back-end of \system are highly decoupled so that the visualization can be used not only with
 Myria but also with other systems such as Spark or Hadoop. In
@@
 \system's user-interface (UI) is embedded into Myria's web front-end.
-We build \system's UI using D3 to delive an interactive visualization experience to the user.
+We build \system's UI using D3\cite{2011-d3} to deliver an interactive visualization experience to the user.
 D3 is a JavaScript framework for data visualziation on the web.

### 2014-03-17 final-report/paper.tex
Focus and context reference
Added fragment view

--- 2014-03-17 final-report/paper.tex
@@
 We propose a visualization system for understanding and exploring query execution and data movement in a distributed database management system (DDBMS).
-Our tool provides insight into: (1) data flow between query fragments and between workers, (2) query execution and operator dependencies, (3) cluster utilization, (4) network utilization.
-Our tool, which we call \system, helps developers understand and improve query execution, and provides insight into common problems such as data skew or performance bottlenecks.
+Our tool will provide insight into: (1) data flow between query fragments and between workers, (2) query execution and operator dependencies, (3) cluster utilization, (4) network utilization.
+Our tool, which we call \system, will help developers understand and improve query execution, and will provide insight into common problems such as data skew or performance bottlenecks.
 
@@
 First, events could be logged to files using a standard logging tool and then collect the files on one node using remote copy.
-Once the logs are on one machine, the logs can be parsed and the data can be transformed into the desired format.
+Once the logs are in one place, they can be parsed and the data transformed into the desired format.
 
@@
   \includegraphics[width=\columnwidth]{images/viz_arch}
-  \caption{Overview of the log collection and transformation architecture.
-Raw event logs are collected on each worker.
-The data is stored in relations.
-To download the data, a query has to be executed.}
+  \caption{Architecture}
   \label{fig:arch}
@@
 \subsection{Front-end}
-
 The Myria web front-end server is written in Python and runs on Google App Engine.
@@
 
-The \fragment view comprises two charts: the aggregation chart and the workers chart.
+The \fragment view comprises two charts: the utilization chart and the
+operators chart. The utilization chart (Fig.~\ref{fig:utilization_chart}) shows the utilization
+of the cluster over time, inside the selected fragment. The chart is used to
+quickly reveal any patterns in the schedule, like tail latency, long periods of
+idleness, etc. We implemented a small brush (at the top of the figure) that
+allows the user to zoom-in and further analyze the problematic area.
+
+\begin{figure}[ht]
+  \includegraphics[width=\columnwidth]{images/utilization_chart}
+  \caption{The utilization chart. }
+  \label{fig:utilization_chart}
+\end{figure}
+
+The utilization chart also implements a brush that allows the user to further
+analyze the query execution by displaying the schedule of each of the worker
+nodes, in the operators chart (Fig.~\ref{fig:operators_chart}). The per-worker
+schedule shows what operators the worker executed and for how long.
+
+\begin{figure}[ht]
+  \includegraphics[width=\columnwidth]{images/operators_chart}
+  \caption{The operators chart.}
+  \label{fig:operators_chart}
+\end{figure}
 
@@
 
-A focus of the current implementation is on scalability to many workers and long running queries.
-More aggregations have to be done by the database itself.
-Most importantly, in the fragment visualization that shows spans for all events we should limit the data that is being downloaded and rendered to what the user can see.
-With the architecture as explained in Section~\ref{sec:back}, we are able to express these filter conditions as queries.
-
-Besides performance improvements, we plan to integrate X-trace and its visualization to offer the orthogonal view and allow users to trace how tuple batches flow through the operators and between operators.
-Combined with the visualizations described in this paper, it will form a powerful debugging tool that will help us to better understand and improve query execution.
+% Dom
 

### 2014-03-17 final-report/paper.tex

--- 2014-03-17 final-report/paper.tex
@@
 D3 is a JavaScript framework for data visualziation on the web.
-We used visualization techniques such as \emph{focus+context} that possible to implement conveniently using D3.
+We used visualization techniques such as \emph{focus+context}\cite{furnas1986generalized} that possible to implement conveniently using D3.
 

### 2014-03-17 final-report/paper.tex
Skew and no skew figures.

--- 2014-03-17 final-report/paper.tex
@@
 \usepackage{xspace}
-
 \usepackage{graphicx}
+\usepackage{caption}
+\usepackage{subcaption}
 
@@
 The \fragment view comprises two charts: the aggregation chart and the workers chart.
+
+\begin{figure}[ht]
+  \centering
+  \begin{subfigure}[b]{0.49\columnwidth}
+    \includegraphics[width=\columnwidth]{images/skew}
+    \caption{Data skew.}
+    \label{fig:skew}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\columnwidth}
+    \includegraphics[width=\columnwidth]{images/no-skew}
+    \caption{No data skew.}
+    \label{fig:no-skew}
+  \end{subfigure}
+  \caption{The network view shows a distinctive pattern when there is data skew. Figure~\ref{fig:skew} shows that most tuples are sent to worker 1 and 5 (origin: x-axis, destination: y-axis).}
+  \label{fig:network}
+\end{figure}
 

### 2014-03-17 final-report/paper.tex
Skew

@@
 % why is a visualization a good way to answer the questions
+% so far the only performance debug mechanism is runtime
 
@@
   \end{subfigure}
-  \caption{The network view shows a distinctive pattern when there is data skew. Figure~\ref{fig:skew} shows that most tuples are sent to worker 1 and 5 (origin: x-axis, destination: y-axis).}
+  \caption{The network view shows a distinctive pattern when there is data skew. Figure~\ref{fig:skew} shows that most tuples are sent to worker 1 and 5 (origin: row, destination: column).}
   \label{fig:network}
@@
 
+In this section we will evaluate the different visualizations and elaborate on actual examples where the visualization helped us to find data or execution skew.
+
 % How it helps
@@
 
-\textbf{Discussion:}
+\subsection{Network visualization}
+
+The network visualization at a glance shows how much data was sent between workers.
+Figure~\ref{fig:network} compares the communication between workers.
+In Figure~\ref{fig:skew} one can see that most of the data is sent to worker~1 and then worker~5.
 
@@
 % Adriana
+
+% Capabilities and Limitations
 

### 2014-03-17 final-report/paper.tex
section on graph

--- 2014-03-17 final-report/paper.tex
@@
 \usepackage{xspace}
+
 \usepackage{graphicx}
-\usepackage{caption}
-\usepackage{subcaption}
 
@@
 
-% What is myria, how does it work (look at report from last quarter but shorter), which languages are supported
+% What is myria, how does it work (look at report from last quarter but shorter), which query languages are supported
 % motivation
-% questions (inspiration: https://docs.google.com/a/uw.edu/document/d/1Syn65bu_p6S4RQETaozsleYcRBauQJ7IduYrctS1wrw/edit?pli=1#)
+% questions & goals
 % why is a visualization a good way to answer the questions
+% contributions
 
@@
 We found that the performance of the query might be affected by the following
-factors, which greatly influenced the design of our visualization tool:
+factors, which influenced :
 \begin{itemize}
-   \item \emph{Wrong/unoptimized physical query plan.} We manually analyzed
-several physical query plans and discovered that some queries were poorly
-optimized. This led us to design a view that allows interactive exploration of
-the physical query plan (mapped to a graph of query operators). 
-   \item \emph{Stragglers, tail latency, expensive query operators, poor
-scheduling.} Our targeted systems are \emph{distributed} databases. Therefore,
-as the query is executed in a distributed environment, the query coordinator
-has to wait until all the workers finish executing their assigned tasks. To
-give insight into how the tasks are distributed and executed at each workers,
-we designed a view that shows the cluster utilization over time and at each
-worker.
-   \item \emph{Poor data partitioning, data skew.} Network traffic can cause
-serious delay when executing a distributed query. Due to poor data
-partitioning, workers might need to send unnecessary big amount of data between
-each other. We designed a view that allows the developer/programmer to analyze
-the data traffic generated while executing the query.  
+   \item 
 \end{itemize}
@@
 
-\system's architecture (Fig.~\ref{fig:arch}) consists of a back-end,
-used as a plug-in interface for the distributed database system targeted, and a
-front-end which implements the views described above. The front-end and
-back-end of \system are highly decoupled so that the visualization can be used not only with
-Myria but also with other systems such as Spark or Hadoop. In
-Section~\ref{sec:back} we illustrate how we implemented log collections and
-aggregation in Myria.
+\begin{itemize}
+    \item determine possible causes (data skew, stragglers, etc.) for why a query might be slow and design visualizations to
+help
+    \item we designed three views - physical query plan, fragment view, communication view.
+\end{itemize}
+
+\system's frontend is embedded into Myria's web frontend and provides comprehensive performance views to the user.
+There are 4 views in total that allow the user to visualize query performance details under different angles:
+\begin{itemize}
+    \item Physical Query Plan view: this view allows the user to connect the query she wrote with the physical query plan that the query optimizer generates.
+    \item etc. 
+\end{itemize}
+
+The front and back end of \system are highly decoupled so that it can be used not only with Myria but also with other systems such as Spark or Hadoop.
+In Section~\ref{sec:back} we illustrate how we implemented log collections and aggregation in Myria.
 
@@
 \begin{figure}[ht]
-  \includegraphics[width=\columnwidth]{images/viz_arch}
-  \caption{Overview of the log collection and transformation architecture.
-Raw event logs are collected on each worker.
-The data is stored in relations.
-To download the data, a query has to be executed.}
+  \includegraphics[width=\columnwidth]{viz_arch}
+  \caption{Architecture}
   \label{fig:arch}
@@
 \subsection{Front-end}
-
 The Myria web front-end server is written in Python and runs on Google App Engine.
 \system's user-interface (UI) is embedded into Myria's web front-end.
-We build \system's UI using D3\cite{2011-d3} to deliver an interactive visualization experience to the user.
+We build \system's UI using D3 to delive an interactive visualization experience to the user.
 D3 is a JavaScript framework for data visualziation on the web.
-We used visualization techniques such as \emph{focus+context}\cite{furnas1986generalized} that possible to implement conveniently using D3.
+We used visualization techniques such as \emph{focus+context} that possible to implement conveniently using D3.
 
@@
 
-\system's web UI is divided into two components: (1) the context window which provides a visualization of the \graph, and (2) the performance window which provides some insight into an element selected in the context view.
-The context window contains a view of the \graph, rendered as a graph.
+\system's web UI is divided into two components: (1) the browser panel which provides a visualization of the \graph, and (2) the performance panel which provides some insight into an element selected in the browser view.
+The browser panel contains a view of the \graph, rendered as a graph.
 The user can navigate the \graph by expanding query fragments into the operators that compose it.
-The user can chose to select a fragment of interest which will render a \fragment visualization for the selected fragment in the performance window.
-Alternatively, the user can select an fragment-to-fragment edge in the \graph thus rendering a \network visualization for the selected edge in the performance window.
-If no elements are selected in the \graph view, an \overall visualization is rendered in the performance window, displaying the aggregate worker utilization over time for each fragment.
+The user can chose to select a fragment of interest which will render a \fragment visualization of the selected fragment in the performance panel.
+Alternatively, the user can select an fragment-to-fragment edge in the \graph thus rendering a \network visualization of the selected edge in the performance panel.
+If no elements are selected in the \graph view, an \overall visualization is rendered in the performance panel, displaying the aggregate worker utilization over time for each fragment.
 
@@
 
+The \graph view provides context to the user.
+It provides a visualization of the physical query plan.
+A query plan is an ordered set of steps used to access data in database management systems.
+The physical query plan is represented by a graph where each node represents a query fragment, and each link represents inter-worker communication from one query fragment to the next.
+Each query fragment is a collection of query operators.
+One can think of a query fragment as a job unit that gets scheduled on a cluster of machines.
+Consequently transitioning from one query fragment to another generally incurs all-to-all communication between worker nodes on the cluster.
+
+The user can perform two classes of actions that will render a new visualization in the performance panel:
+\begin{itemize}
+
+    \item \textbf{Empty selection:} by default if no fragments or edges are selected in the \graph view, an \overall visualization is rendered in the performance panel.
+This visualization displays the aggregate worker utilization over time for each fragment and provides cluster utilization data at a glance.
+    \item \textbf{Fragment selection:} when selecting a fragment, a \fragment visualization gets displayed in the performance panel.
+The \fragment visualization provides aggregate cluster utilization information complemented by per-worker task schedules.
+When a fragment is selected, the operands inside of the fragment in the \graph view are color-coded to allow the user to easily match each task in the per-worker schedule in the performance window with the corresponding query operator.
+    \item \textbf{Edge selection:} upon selecting one or more fragment-to-fragment edges, a \network visualization gets displayed, providing aggregate exchange data between cluster nodes.
+
+\end{itemize}
+
+\textbf{Graph Rendering:} We use D3 to render the \graph graph, which allows us to support various interactions and transistions.
+We use GraphViz \cite{graphviz} in the backend to generate graph layout information.
+GraphViz allows us to generate graph layouts optimized for space.
+The layout information is then fed into a D3-based rendering engine.
+
+\textbf{Graph Navigation:} A query plan can be arbitrarily large.
+Thus we offer two mechanisms that facilitate exploration of the graph for the user: (1) expanding/reducing fragments and (2) zooming.
+The first mechanism can reduce the size of the graph by a constant factor by collapsing operators that compose a fragment into a single node.
+The user can click once on a collapsed fragment to expand it, and click once on an expanded fragment to collapse it.
+Expanding or reducing a node can cause large changes in the graph layout because GraphViz optimizes the graph layout to minimize area.
+We implemented transitions to allow the user to track the fragments as they get shuffled around after one fragment gets expanded or reduced.
+The second mechanism that facilitates exploration of the graph for the user is zooming.
+This feature was implemented in D3 and allows the user to quickly jump from one part of the graph to another, by zooming in and out of the view.
+
+\textbf{Tooltip Information:} TODO
+
 \subsubsection{Fragment execution view}
@@
 % Adriana
-
-The \fragment view comprises two charts: the aggregation chart and the workers chart.
-
-\begin{figure}[ht]
-  \centering
-  \begin{subfigure}[b]{0.49\columnwidth}
-    \includegraphics[width=\columnwidth]{images/skew}
-    \caption{Data skew.}
-    \label{fig:skew}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\columnwidth}
-    \includegraphics[width=\columnwidth]{images/no-skew}
-    \caption{No data skew.}
-    \label{fig:no-skew}
-  \end{subfigure}
-  \caption{The network view shows a distinctive pattern when there is data skew. Figure~\ref{fig:skew} shows that most tuples are sent to worker 1 and 5 (origin: x-axis, destination: y-axis).}
-  \label{fig:network}
-\end{figure}
 

### 2014-03-17 final-report/paper.tex

--- 2014-03-17 final-report/paper.tex
@@
 
-The \fragment view comprises two charts: the aggregation chart and the workers chart.
+The \fragment view comprises two charts: the utilization chart and the
+operators chart. The utilization chart (Fig.~\ref{fig:utilization_chart}) shows the utilization
+of the cluster over time, inside the selected fragment. The chart is used to
+quickly reveal any patterns in the schedule, like tail latency, long periods of
+idleness, etc. We implemented a small brush (at the top of the figure) that
+allows the user to zoom-in and further analyze the problematic area.
+
+\begin{figure}[ht]
+  \includegraphics[width=\columnwidth]{images/utilization_chart}
+  \caption{The utilization chart. }
+  \label{fig:utilization_chart}
+\end{figure}
+
+The utilization chart also implements a brush that allows the user to further
+analyze the query execution by displaying the schedule of each of the worker
+nodes, in the operators chart (Fig.~\ref{fig:operators_chart}). The per-worker
+schedule shows what operators the worker executed and for how long.
+
+\begin{figure}[ht]
+  \includegraphics[width=\columnwidth]{images/operators_chart}
+  \caption{The operators chart.}
+  \label{fig:operators_chart}
+\end{figure}
+
+% use the word spans when referring to events with duration (used in dapper)
+
+\subsubsection{Worker communication view}
+
+% Umar
+
+\section{Evaluation}
 
@@
 \end{figure}
-
-% use the word spans when referring to events with duration (used in dapper)
-
-\subsubsection{Worker communication view}
-
-% Umar
-
-\section{Evaluation}
 

### 2014-03-17 final-report/paper.tex
Typo

@@
 
-\system's architecture (Fig.~\ref{fig:arch}) consists of a back-end,
+\system's architecture (Figure~\ref{fig:arch}) consists of a back-end,
 used as a plug-in interface for the distributed database system targeted, and a
@@
 We build \system's UI using D3\cite{2011-d3} to deliver an interactive visualization experience to the user.
-D3 is a JavaScript framework for data visualziation on the web.
+D3 is a JavaScript framework for data visualization on the web.
 We used visualization techniques such as \emph{focus+context}\cite{furnas1986generalized} that possible to implement conveniently using D3.

### 2014-03-17 final-report/paper.tex

@@
 
-% What is myria, how does it work (look at report from last quarter but shorter), which languages are supported
+% What is myria, how does it work (look at report from last quarter but shorter), which query languages are supported
 % motivation
@@
 % so far the only performance debug mechanism is runtime
+% contributions
 
@@
 % Thierry
+
+The \graph view provides context to the user.
+It provides a visualization of the physical query plan.
+A query plan is an ordered set of steps used to access data in database management systems.
+The physical query plan is represented by a graph where each node represents a query fragment, and each link represents inter-worker communication from one query fragment to the next.
+Each query fragment is a collection of query operators.
+One can think of a query fragment as a job unit that gets scheduled on a cluster of machines.
+Consequently transitioning from one query fragment to another generally incurs all-to-all communication between worker nodes on the cluster.
+
+The user can perform two classes of actions that will render a new visualization in the performance panel:
+\begin{itemize}
+
+    \item \textbf{Empty selection:} by default if no fragments or edges are selected in the \graph view, an \overall visualization is rendered in the performance panel.
+This visualization displays the aggregate worker utilization over time for each fragment and provides cluster utilization data at a glance.
+    \item \textbf{Fragment selection:} when selecting a fragment, a \fragment visualization gets displayed in the performance panel.
+The \fragment visualization provides aggregate cluster utilization information complemented by per-worker task schedules.
+When a fragment is selected, the operands inside of the fragment in the \graph view are color-coded to allow the user to easily match each task in the per-worker schedule in the performance window with the corresponding query operator.
+    \item \textbf{Edge selection:} upon selecting one or more fragment-to-fragment edges, a \network visualization gets displayed, providing aggregate exchange data between cluster nodes.
+
+\end{itemize}
+
+\textbf{Graph Rendering:} We use D3 to render the \graph graph, which allows us to support various interactions and transistions.
+We use GraphViz \cite{graphviz} in the backend to generate graph layout information.
+GraphViz allows us to generate graph layouts optimized for space.
+The layout information is then fed into a D3-based rendering engine.
+
+\textbf{Graph Navigation:} A query plan can be arbitrarily large.
+Thus we offer two mechanisms that facilitate exploration of the graph for the user: (1) expanding/reducing fragments and (2) zooming.
+The first mechanism can reduce the size of the graph by a constant factor by collapsing operators that compose a fragment into a single node.
+The user can click once on a collapsed fragment to expand it, and click once on an expanded fragment to collapse it.
+Expanding or reducing a node can cause large changes in the graph layout because GraphViz optimizes the graph layout to minimize area.
+We implemented transitions to allow the user to track the fragments as they get shuffled around after one fragment gets expanded or reduced.
+The second mechanism that facilitates exploration of the graph for the user is zooming.
+This feature was implemented in D3 and allows the user to quickly jump from one part of the graph to another, by zooming in and out of the view.
+
+\textbf{Tooltip Information:} TODO
 

### 2014-03-17 final-report/paper.tex
graph view changes

@@
 
-The \graph view provides context to the user.
-It provides a visualization of the physical query plan.
+The \graph view allows the user to selectively browse different performance visualizations tied to a given query execution.
+The \graph view provides as the name indicates, a visualization of the physical query plan.
 A query plan is an ordered set of steps used to access data in database management systems.
@@
 One can think of a query fragment as a job unit that gets scheduled on a cluster of machines.
-Consequently transitioning from one query fragment to another generally incurs all-to-all communication between worker nodes on the cluster.
+Consequently transitioning from one query fragment to another generally leads to all-to-all communication between worker nodes on the cluster.
 
-The user can perform two classes of actions that will render a new visualization in the performance panel:
+The user can perform three classes of actions that will render a new visualization in the performance panel:
 \begin{itemize}
@@
 This visualization displays the aggregate worker utilization over time for each fragment and provides cluster utilization data at a glance.
-    \item \textbf{Fragment selection:} when selecting a fragment, a \fragment visualization gets displayed in the performance panel.
+    \item \textbf{Fragment selection:} when selecting a fragment, a \fragment visualization gets rendered in the performance panel.
 The \fragment visualization provides aggregate cluster utilization information complemented by per-worker task schedules.
-When a fragment is selected, the operands inside of the fragment in the \graph view are color-coded to allow the user to easily match each task in the per-worker schedule in the performance window with the corresponding query operator.
-    \item \textbf{Edge selection:} upon selecting one or more fragment-to-fragment edges, a \network visualization gets displayed, providing aggregate exchange data between cluster nodes.
+When a fragment is selected, the operands inside of the fragment in the \graph view are color-coded to allow the user to easily match each task in the per-worker task schedule in the performance window with the corresponding query operator in the browser window.
+    \item \textbf{Edge selection:} upon selecting one or more fragment-to-fragment edges, a \network visualization gets displayed, providing information on inter-worker communication from one query fragment to the next.
 
@@
 We use GraphViz \cite{graphviz} in the backend to generate graph layout information.
-GraphViz allows us to generate graph layouts optimized for space.
-The layout information is then fed into a D3-based rendering engine.
+GraphViz allows us to generate graph layouts that are optimized to minimize area footprint.
+The layout information is then fed into a D3-based rendering engine which supports various interactions and animations.
 
@@
 Thus we offer two mechanisms that facilitate exploration of the graph for the user: (1) expanding/reducing fragments and (2) zooming.
-The first mechanism can reduce the size of the graph by a constant factor by collapsing operators that compose a fragment into a single node.
+The first mechanism can reduce the size of a graph by a constant factor by collapsing operators that compose a fragment into a single fragment node.
 The user can click once on a collapsed fragment to expand it, and click once on an expanded fragment to collapse it.
-Expanding or reducing a node can cause large changes in the graph layout because GraphViz optimizes the graph layout to minimize area.
-We implemented transitions to allow the user to track the fragments as they get shuffled around after one fragment gets expanded or reduced.
+Expanding or reducing a node can cause large changes in the graph layout as GraphViz changes the graph layout to minimize overall area.
+To address these layout changes, we implemented transitions to allow the user to track the fragments as those get reshuffled.
 The second mechanism that facilitates exploration of the graph for the user is zooming.
-This feature was implemented in D3 and allows the user to quickly jump from one part of the graph to another, by zooming in and out of the view.
+This feature was implemented in D3 and allows the user to quickly jump from one part of the graph to the next, by zooming in and out of the graph view.
 

### 2014-03-18 final-report/paper.tex
Graphviz reference
Remove trailing spaces
Typo

@@
 
-\textbf{Graph Rendering:} We use D3 to render the \graph graph, which allows us to support various interactions and transistions.
-We use GraphViz \cite{graphviz} in the backend to generate graph layout information.
+\textbf{Graph Rendering:} We use D3 to render the \graph graph, which allows us to support various interactions and transitions.
+We use GraphViz \cite{Ellson01graphviz} in the backend to generate graph layout information.
 GraphViz allows us to generate graph layouts that are optimized to minimize area footprint.

### 2014-03-18 final-report/paper.tex
results section

@@
 \usepackage{subcaption}
+\usepackage{listings}
 
@@
 \author{\alignauthor Umar Javed, Thierry Moreau, Dominik Moritz, Adriana Szekeres \\
-\affaddr{Dept. of Computer Science, University of Washington} \\ \affaddr{ Seattle, Washington, USA} \\
+\affaddr{Dept. of Computer Science, University of Washington} \\ \affaddr{Seattle, Washington, USA} \\
 \email{\texttt{\{ujaved, moreau, domoritz, aaasz\}@cs.washington.edu}}
@@
     \item motivation: help developers/programmers dig what the problems might be
+    \item Questions: find data/execution skew, identify performance bottlenecks, visualize data flow in the system (elaborate on each point)...
 \end{itemize}
@@
 
-The Myria web front-end server is written in Python and runs on Google App Engine.
+The Myria web front-end server is written in Python and runs on Google App Engine\footnote{\url{https://developers.google.com/appengine/}}.
 \system's user-interface (UI) is embedded into Myria's web front-end.
@@
 
+% Figures
+% 1: Graph with back-edge
+% 2: Large Graph expanded vs. large Graph reduced vs. large graph expanded -1
+
 The \graph view allows the user to selectively browse different performance visualizations tied to a given query execution.
@@
 
-\textbf{Graph Rendering:} We use D3 to render the \graph graph, which allows us to support various interactions and transitions.
+\textbf{Graph Rendering:} We use D3 \cite{d3} to render the \graph graph, which allows us to support various interactions and transistions.
 We use GraphViz \cite{Ellson01graphviz} in the backend to generate graph layout information.
@@
 
+We used \system to examine real world database queries that ran on Myria.
+We base our analysis on the different visualizations that were generated by \system and present examples where the \system helped us better identify (1) data skew, (2) execution skew, (3) performance bottlenecks.
+
+We explore a simple join query we performed on one million tuples extracted from Twitter data. This query joins Twitter followers and followees by user ID. The query written in SQL is expressed below.
+
+\begin{lstlisting}
+SELECT *
+FROM twitter S, twitter R
+WHERE S.follower = R.followee
+\end{lstlisting}
+
+The resulting \graph visualization is displayed in ~\ref{fig:TODO}.
+The user will navigate the \graph visualization in the browser panel in the next case studies to identify several performance symptoms.
+
+% How it helps
+% Examples
+
+% Thierry + Dom
+
+\subsection{Case 1: Identifying Performance Bottlenecks using the \overall View}
+
+%dom
+
+\subsection{Case 1: Identifying Data Skew using the \network View}
+
 \begin{figure}[ht]
@@
 \end{figure}
-
-In this section we will evaluate the different visualizations and elaborate on actual examples where the visualization helped us to find data or execution skew.
-
-% How it helps
-% Examples
-
-% Thierry + Dom
-
-\subsection{Network visualization}
 

### 2014-03-18 final-report/paper.tex
Short paragraph about overview chart

--- 2014-03-18 final-report/paper.tex
@@
 \newcommand*{\fragment}{Fragment Execution\xspace}
+\newcommand*{\overview}{Fragment Overview\xspace}
 \newcommand*{\network}{Worker Communication\xspace}
@@
 % contributions
+% explain difference between execution and data skew
 
@@
 
-\subsubsection{Overview over all fragments}
-\label{sec:fragments}
-
-% Adriana
-
-% Write about small multiples view
-
 \subsubsection{Physical query plan view}
@@
 \subsubsection{Fragment execution view}
+\label{sec:fragment}
 
@@
 The \fragment view comprises two charts: the utilization chart and the
-operators chart. The utilization chart (Fig.~\ref{fig:utilization_chart}) shows the utilization
-of the cluster over time, inside the selected fragment. The chart is used to
+operators chart. The utilization chart (Fig.~\ref{fig:utilization_chart}) shows on
+how many workers a fragment is running over time. The chart is used to
 quickly reveal any patterns in the schedule, like tail latency, long periods of
@@
 % use the word spans when referring to events with duration (used in dapper)
+
+\subsubsection{Overview over all fragments}
+\label{sec:fragments}
+
+% Dom
+
+The \overview shows small multiples of the utilization chart as described in the previous section.
+Execution skew can be spotted in this view. Furthermore, the user can compare the fragments and find
+correlations between the execution of different fragments. This view helps the user navigate to the
+right fragment and is not intended to offer further inside into the causes or performance problems.
+
+% Write about small multiples view
 

### 2014-03-18 final-report/paper.tex
Add todos

@@
 % explain difference between execution and data skew
+
+\todo{write me}
 
@@
 
-\textbf{Tooltip Information:} TODO
+\textbf{Tooltip Information:} \todo{write me}
 
@@
 
+\todo{write me}
+
 \section{Evaluation}
@@
 
-We explore a simple join query we performed on one million tuples extracted from Twitter data. This query joins Twitter followers and followees by user ID. The query written in SQL is expressed below.
+We explore a simple join query we performed on one million tuples extracted from the Twitter user graph.
+This query joins Twitter followers and followees by user ID.
+The query can be written in SQL as expressed below.
 
@@
 
-The resulting \graph visualization is displayed in ~\ref{fig:TODO}.
+The resulting \graph visualization is displayed in Figure~\ref{fig:TODO}.
 The user will navigate the \graph visualization in the browser panel in the next case studies to identify several performance symptoms.
@@
 %dom
+
+\todo{write me}
 
@@
 
+\todo{write me}
+
 \section{Future Work}

### 2014-03-18 final-report/paper.tex
Refactor

@@
 \newcommand*{\fragment}{Fragment Execution\xspace}
-\newcommand*{\overview}{Fragment Overview\xspace}
 \newcommand*{\network}{Worker Communication\xspace}
-\newcommand*{\overall}{Fragment Overall\xspace}
+\newcommand*{\overall}{Fragment Overview\xspace}
 
@@
 
-The \overview shows small multiples of the utilization chart as described in the previous section.
+The \overall shows small multiples of the utilization chart as described in the previous section.
 Execution skew can be spotted in this view. Furthermore, the user can compare the fragments and find
@@
 
-\subsection{Case 1: Identifying Performance Bottlenecks using the \overall View}
+\subsection{Case 1: Identifying Performance Bottlenecks using the \overall}
 

### 2014-03-18 final-report/paper.tex
results

--- 2014-03-18 final-report/paper.tex
@@
 \newcommand*{\fragment}{Fragment Execution\xspace}
-\newcommand*{\overview}{Fragment Overview\xspace}
 \newcommand*{\network}{Worker Communication\xspace}
@@
 % contributions
-% explain difference between execution and data skew
 
@@
 
+\subsubsection{Overview over all fragments}
+\label{sec:fragments}
+
+% Adriana
+
+% Write about small multiples view
+
 \subsubsection{Physical query plan view}
@@
 \subsubsection{Fragment execution view}
-\label{sec:fragment}
 
@@
 The \fragment view comprises two charts: the utilization chart and the
-operators chart. The utilization chart (Fig.~\ref{fig:utilization_chart}) shows on
-how many workers a fragment is running over time. The chart is used to
+operators chart. The utilization chart (Fig.~\ref{fig:utilization_chart}) shows the utilization
+of the cluster over time, inside the selected fragment. The chart is used to
 quickly reveal any patterns in the schedule, like tail latency, long periods of
@@
 % use the word spans when referring to events with duration (used in dapper)
-
-\subsubsection{Overview over all fragments}
-\label{sec:fragments}
-
-% Dom
-
-The \overview shows small multiples of the utilization chart as described in the previous section.
-Execution skew can be spotted in this view. Furthermore, the user can compare the fragments and find
-correlations between the execution of different fragments. This view helps the user navigate to the
-right fragment and is not intended to offer further inside into the causes or performance problems.
-
-% Write about small multiples view
 
@@
 
-\subsection{Case 1: Identifying Data Skew using the \network View}
+\subsection{Case 2: Identifying Data Skew using the \network View}
 
@@
 
-The network visualization at a glance shows how much data was sent between workers.
-Figure~\ref{fig:network} compares the communication between workers.
-In Figure~\ref{fig:skew} one can see that most of the data is sent to worker~1 and then worker~5.
+The \network visualization at a glance shows how much data was sent between workers by one query fragment to the next.
+The two \network views in Figure~\ref{fig:no-skew} and Figure~\ref{fig:skew} were obtained by selecting the Frag0$\rightarrow$Frag2 and Frag1$\rightarrow$Frag2 edges respectively in the \graph view.
+Figure~\ref{fig:network} displays aggregate data communication between workers.
+In Figure~\ref{fig:no-skew} the amount of data sent from each worker to every other worker is balanced as opposed to Figure~\ref{fig:skew} where one can see that the volume of data sent to worker~1 and worker~5 outweights data sent to the other workers in the system.
+This apparent data skew could be caused by partitioning issues.
+
+\subsection{Case 3: Identifying Execution Skew using the \fragment View}
+The \fragment visualization allows the user to dive down into query execution schedules at the worker level.
 

### 2014-03-18 final-report/paper.tex

--- 2014-03-18 final-report/paper.tex
@@
 
-\todo{write me}
-
-\subsection{Case 1: Identifying Data Skew using the \network View}
+\subsection{Case 2: Identifying Data Skew using the \network View}
 

### 2014-03-18 final-report/paper.tex
Bug

--- 2014-03-18 final-report/paper.tex
@@
 \usepackage{graphicx}
+
 \usepackage{caption}
+% Bug in my version of the caption package
+\DeclareCaptionType{copyrightbox}
 \usepackage{subcaption}
-\usepackage{listings}
 
@@
 \newcommand*{\network}{Worker Communication\xspace}
-\newcommand*{\overall}{Fragment Overview\xspace}
+\newcommand*{\overall}{Fragment Overall\xspace}
 
@@
 \author{\alignauthor Umar Javed, Thierry Moreau, Dominik Moritz, Adriana Szekeres \\
-\affaddr{Dept. of Computer Science, University of Washington} \\ \affaddr{Seattle, Washington, USA} \\
+\affaddr{Dept. of Computer Science, University of Washington} \\ \affaddr{ Seattle, Washington, USA} \\
 \email{\texttt{\{ujaved, moreau, domoritz, aaasz\}@cs.washington.edu}}
@@
 % contributions
-% explain difference between execution and data skew
-
-\todo{write me}
 
@@
     \item motivation: help developers/programmers dig what the problems might be
-    \item Questions: find data/execution skew, identify performance bottlenecks, visualize data flow in the system (elaborate on each point)...
 \end{itemize}
@@
 
-The Myria web front-end server is written in Python and runs on Google App Engine\footnote{\url{https://developers.google.com/appengine/}}.
+The Myria web front-end server is written in Python and runs on Google App Engine.
 \system's user-interface (UI) is embedded into Myria's web front-end.
@@
 % Thierry
-
-% Figures
-% 1: Graph with back-edge
-% 2: Large Graph expanded vs. large Graph reduced vs. large graph expanded -1
 
@@
 
-\textbf{Graph Rendering:} We use D3 \cite{d3} to render the \graph graph, which allows us to support various interactions and transistions.
+\textbf{Graph Rendering:} We use D3 to render the \graph graph, which allows us to support various interactions and transistions.
 We use GraphViz \cite{Ellson01graphviz} in the backend to generate graph layout information.
@@
 
-\textbf{Tooltip Information:} \todo{write me}
+\textbf{Tooltip Information:} TODO
 
 \subsubsection{Fragment execution view}
-\label{sec:fragment}
 
@@
 
-\subsubsection{Overview over all fragments}
-\label{sec:fragments}
-
-% Dom
-
-The \overall shows small multiples of the utilization chart as described in the previous section.
-Execution skew can be spotted in this view. Furthermore, the user can compare the fragments and find
-correlations between the execution of different fragments. This view helps the user navigate to the
-right fragment and is not intended to offer further inside into the causes or performance problems.
-
-% Write about small multiples view
-
 \subsubsection{Worker communication view}
@@
 
-\todo{write me}
-
 \section{Evaluation}
-
-We used \system to examine real world database queries that ran on Myria.
-We base our analysis on the different visualizations that were generated by \system and present examples where the \system helped us better identify (1) data skew, (2) execution skew, (3) performance bottlenecks.
-
-We explore a simple join query we performed on one million tuples extracted from the Twitter user graph.
-This query joins Twitter followers and followees by user ID.
-The query can be written in SQL as expressed below.
-
-\begin{lstlisting}
-SELECT *
-FROM twitter S, twitter R
-WHERE S.follower = R.followee
-\end{lstlisting}
-
-The resulting \graph visualization is displayed in Figure~\ref{fig:TODO}.
-The user will navigate the \graph visualization in the browser panel in the next case studies to identify several performance symptoms.
-
-% How it helps
-% Examples
-
-% Thierry + Dom
-
-\subsection{Case 1: Identifying Performance Bottlenecks using the \overall}
-
-%dom
-
-\todo{write me}
-
-\subsection{Case 1: Identifying Data Skew using the \network View}
 
@@
 
+In this section we will evaluate the different visualizations and elaborate on actual examples where the visualization helped us to find data or execution skew.
+
+% How it helps
+% Examples
+
+% Thierry + Dom
+
+\subsection{Network visualization}
+
 The network visualization at a glance shows how much data was sent between workers.
@@
 % Capabilities and Limitations
-
-\todo{write me}
 

### 2014-03-18 final-report/paper.tex
added figures, discussion etc.

--- 2014-03-18 final-report/paper.tex
@@
 
-The \graph view allows the user to selectively browse different performance visualizations tied to a given query execution.
+\begin{figure}[ht]
+  \centering
+  \begin{subfigure}[b]{0.49\columnwidth}
+    \includegraphics[width=0.8\columnwidth]{images/graph_collapsed}
+    \caption{Large query plan collapsed.}
+    \label{fig:graph}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\columnwidth}
+    \includegraphics[width=0.8\columnwidth]{images/graph_expanded}
+    \caption{Expanding Fragment 2 reveals its operators.}
+    \label{fig:graph_expanded}
+  \end{subfigure}
+  \caption{The \graph view is used by \system to help the user browse performance visualizations.}
+  \label{fig:graph}
+\end{figure}
+
+The \graph view as pictured in Figure~\ref{fig:graph} allows the user to selectively browse different performance visualizations tied to a given query execution.
 The \graph view provides as the name indicates, a visualization of the physical query plan.
@@
 GraphViz allows us to generate graph layouts that are optimized to minimize area footprint.
-The layout information is then fed into a D3-based rendering engine which supports various interactions and animations.
+The layout information is then fed into a D3-based rendering engine which supports various interactions and animations techniques.
 
@@
 The first mechanism can reduce the size of a graph by a constant factor by collapsing operators that compose a fragment into a single fragment node.
-The user can click once on a collapsed fragment to expand it, and click once on an expanded fragment to collapse it.
+The user can click once on a collapsed fragment to expand, and click once on an expanded fragment to collapse it.
 Expanding or reducing a node can cause large changes in the graph layout as GraphViz changes the graph layout to minimize overall area.
+This graph reshuffeling is illustrated in Figure~\ref{graph} where expanding Frag2 causes the layout to change.
 To address these layout changes, we implemented transitions to allow the user to track the fragments as those get reshuffled.
@@
 
-The resulting \graph visualization is displayed in Figure~\ref{fig:TODO}.
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=0.4\columnwidth]{images/graph_join}
+  \caption{This view shows the \graph for a join.}
+  \label{fig:join}
+\end{figure}
+
+The resulting \graph visualization is displayed in Figure~\ref{fig:join}.
 The user will navigate the \graph visualization in the browser panel in the next case studies to identify several performance symptoms.
@@
 \subsection{Case 3: Identifying Execution Skew using the \fragment View}
-The \fragment visualization allows the user to dive down into query execution schedules at the worker level.
+
+\begin{figure}[ht]
+  \includegraphics[width=\columnwidth]{images/execution_skew}
+  \caption{This \fragment view shows heavy execution skew on worker~5.}
+  \label{fig:exec_skew}
+\end{figure}
+
+The \fragment visualization allows the user to dive down into query execution details at the worker level.
+The \fragment view in Figure~\ref{fig:exec_skew} was obtained by selecting Frag2, which is the fragment that performs the join, in the \graph view.
+The user can observe at a glance that certain workers produce much more data as a result of the join execution.
+Particularly, we saw in Case 2 Figure~\ref{fig:skew}, that tuples sent to Fragment 2 from Fragment 1 were heavily skewed towards being sent to worker~1 and worker~5.
+In Figure~\ref{fig:exec_skew}, the user can now see that as a result of the join, worker~5 ends up producing most tuples that have to be written back to disk.
+The color scheme used for each worker schedules in the \fragment view is color-coded to match the operators in Figure~\ref{fig:join}.
+Consequently, one can quickly observe in the \fragment view that worker~5 is busy sending join results from t=2.5s to t=5.0s, which differes from the rest of the workers.
+Using similar reasoning, one can conclude that worker 5 presents a performance bottleneck in the execution of the query on Myria.
+Perhaps the user can modify the way the input data set gets partitioned.
+Alternatively she could change the number of workers allocated for the job.
+
+\subsection{Discussion}
+
+All in all, \system allows the user to get much more insight into the execution of a query than what was available before, i.e.
+text-based performance logs, and runtimes.
+Being able to navigate performance data using a simple interface can indeed facilitate the task of a developper who is designing Myria's back end, or a Myria user who is writing complex queries to process large amounts of data.
+That said, \system can only help the user visualize query execution performance, but won't necessarily point the user to what she needs to do to improve performance of her query execution.
 

### 2014-03-18 final-report/paper.tex
part of intro

--- 2014-03-18 final-report/paper.tex
@@
 \newcommand*{\fragment}{Fragment Execution\xspace}
+\newcommand*{\overview}{Fragment Overview\xspace}
 \newcommand*{\network}{Worker Communication\xspace}
-\newcommand*{\overall}{Fragment Overview\xspace}
+\newcommand*{\overall}{Fragment Overall\xspace}
 
@@
 \section{Introduction}
+
+Database query performance is often unpredictable, specially when the DBMS is distributed. This requires database 
+administrators to spend significant amounts of time and effort isolating and debugging the root cause of performance 
+problems. These may include skewed data partitioning or run-time issues, such as resource contention or 
+adverse network conditions.
+
+Our goal is to build a visualization tool designed to help a database administrator understand the data flow and run-time
+performance during query execution in a Distributed Database Management System (DDBMS). Using our tool, the
+database programmer or administrator could quickly discover run-time bottlenecks as well as problematic data skew issues, 
+saving them significant guesswork and programming effort to identify the appropriate logs to analyze.
 
@@
 
-The \overall shows small multiples of the utilization chart as described in the previous section.
+The \overview shows small multiples of the utilization chart as described in the previous section.
 Execution skew can be spotted in this view. Furthermore, the user can compare the fragments and find
@@
 
-\subsection{Case 1: Identifying Performance Bottlenecks using the \overall}
+\subsection{Case 1: Identifying Performance Bottlenecks using the \overall View}
 
@@
 
-\subsection{Case 2: Identifying Data Skew using the \network View}
+\todo{write me}
+
+\subsection{Case 1: Identifying Data Skew using the \network View}
 
@@
 
-The \network visualization at a glance shows how much data was sent between workers by one query fragment to the next.
-The two \network views in Figure~\ref{fig:no-skew} and Figure~\ref{fig:skew} were obtained by selecting the Frag0$\rightarrow$Frag2 and Frag1$\rightarrow$Frag2 edges respectively in the \graph view.
-Figure~\ref{fig:network} displays aggregate data communication between workers.
-In Figure~\ref{fig:no-skew} the amount of data sent from each worker to every other worker is balanced as opposed to Figure~\ref{fig:skew} where one can see that the volume of data sent to worker~1 and worker~5 outweights data sent to the other workers in the system.
-This apparent data skew could be caused by partitioning issues.
-
-\subsection{Case 3: Identifying Execution Skew using the \fragment View}
-The \fragment visualization allows the user to dive down into query execution schedules at the worker level.
+The network visualization at a glance shows how much data was sent between workers.
+Figure~\ref{fig:network} compares the communication between workers.
+In Figure~\ref{fig:skew} one can see that most of the data is sent to worker~1 and then worker~5.
 

### 2014-03-18 final-report/paper.tex
more intro

@@
 Myria aims towards building a distributed database platform to provide \emph{big data management and analytics as a service} primarily for scientific applications.
-
 The proposed visualizations can easily be applied to other DDBMS as well (e.g. Spark, Hadoop).
@@
 database programmer or administrator could quickly discover run-time bottlenecks as well as problematic data skew issues, 
-saving them significant guesswork and programming effort to identify the appropriate logs to analyze.
+saving them significant guesswork and programming effort to identify the appropriate logs to analyze. Our visualization tool
+is called \system, which uses profiling logs from Myria \cite{myria} as the prototype DDMS. Myria is a distributed 
+database platform that provides \emph{big data management and analytics as a service} primarily for scientific applications.
+Myria takes queries in high-level query languages, such as SQL, translating them into physical query plans and finally executing
+them in parallel. Currently
 
@@
 % explain difference between execution and data skew
-
-\todo{write me}
 

### 2014-03-18 final-report/paper.tex
more info

@@
 Myria takes queries in high-level query languages, such as SQL, translating them into physical query plans and finally executing
-them in parallel. Currently
+them in parallel. Currently, Myria does not offer a visualization tool designed to show the physical query plan, job execution and network 
+communication between distributed workers. \system allows the programmer/database administrator to visualize the following aspects
+of query execution.
+
+\begin{itemize}
+
+\item An interactive visualization of the physical query plan. The physical query plan in Myria is composed of
+a DAG of fragments, where each DAG link represents communication between workers, possibly over the network. 
+\system displays this DAG in an interactive fashion with clickable elements for fragment nodes and communication links. 
+This DAG also represents a high-level data flow in the DDBMS.
+
+\end{itemize}
 
@@
 
-\begin{itemize}
-    \item problem: hard to learn why a query is slow
-    \item motivation: help developers/programmers dig what the problems might be
-    \item Questions: find data/execution skew, identify performance bottlenecks, visualize data flow in the system (elaborate on each point)...
-\end{itemize}
+% \item Questions: find data/execution skew, identify performance bottlenecks, visualize data flow in the system (elaborate on each point)...
 

### 2014-03-19 final-report/paper.tex
more intro

@@
 saving them significant guesswork and programming effort to identify the appropriate logs to analyze. Our visualization tool
-is called \system, which uses profiling logs from Myria \cite{myria} as the prototype DDMS. Myria is a distributed 
+is called \system, which uses profiling logs from \emph{Myria} \cite{myria} as the prototype DDMS. Myria is a distributed 
 database platform that provides \emph{big data management and analytics as a service} primarily for scientific applications.
@@
 This DAG also represents a high-level data flow in the DDBMS.
+
+\item A view for cluster utilization and operator performance for each \emph{Myria} segment. Each segment is composed of
+multiple sequential database operators running on the same worker. Upon clicking on a collapsed segment node in the
+query plan DAG, the node expands to display the operators comprising the segment and renders the fragment execution view.
+The fragment execution view visualizes the utilization of the cluster by showing whether each worker is busy working 
+on the selected fragment. It also breaks down the work done by each worker into operators composing the fragment.
 

### 2014-03-19 final-report/paper.tex
more intro

@@
 on the selected fragment. It also breaks down the work done by each worker into operators composing the fragment.
+
+\item A view for communication between workers. A directed link between two fragments in the physical query plan DAG
+represents source workers sending tuples to destination workers. Upon clicking on a link, the visualization renders the
+worker communication view. This shows the aggregate number of tuples exchanged between workers as well as a time series
+charting the number of tuples exchanged over time.
 

### 2014-03-19 final-report/paper.tex
Defined what a query fragment is. Small fixes to the abstract and intro..

@@
 \usepackage{graphicx}
+
 \usepackage{caption}
+% Bug in my version of the caption package
+\DeclareCaptionType{copyrightbox}
 \usepackage{subcaption}
@@
 
-We propose a visualization system for understanding and exploring query execution and data movement in a distributed database management system (DDBMS).
-Our tool provides insight into: (1) data flow between query fragments and between workers, (2) query execution and operator dependencies, (3) cluster utilization, (4) network utilization.
-Our tool, which we call \system, helps developers understand and improve query execution, and provides insight into common problems such as data skew or performance bottlenecks.
+We propose a visualization system to explore and understand query
+execution and data movement in a distributed database management system
+(DDBMS). Our tool provides insight into: (1) data flow between query segments
+and between workers, (2) query execution and operator dependencies, (3) cluster
+utilization, (4) network utilization. Our tool, \system, helps
+developers understand and improve query execution, and provides insight into
+common problems such as data skew or performance bottlenecks.
 
-In particular, \system is built to inspect query execution in Myria\footnote{\url{http://myria.cs.washington.edu/}}, a distributed big data management system currently being developed in the UW CSE database group.
-Myria aims towards building a distributed database platform to provide \emph{big data management and analytics as a service} primarily for scientific applications.
-The proposed visualizations can easily be applied to other DDBMS as well (e.g. Spark, Hadoop).
+In particular, \system is built to inspect query execution in
+Myria\footnote{\url{http://myria.cs.washington.edu/}}, a distributed big data
+management system currently being developed in the UW CSE database group. Myria
+aims towards building a distributed database platform to provide \emph{big data
+management and analytics as a service} primarily for scientific applications.
+The proposed visualizations can easily be applied to other DDBMS as well (e.g.
+Spark, Hadoop).
 
@@
 
-Database query performance is often unpredictable, specially when the DBMS is distributed. This requires database 
-administrators to spend significant amounts of time and effort isolating and debugging the root cause of performance 
-problems. These may include skewed data partitioning or run-time issues, such as resource contention or 
-adverse network conditions.
+Database query performance is often unpredictable, especially when the DBMS is
+distributed. This requires programmers or database administrators to spend
+significant amounts of time and effort isolating and debugging the root cause
+of performance problems. These may include skewed data partitioning or run-time
+issues, such as resource contention or adverse network conditions.
 
@@
 
-\item An interactive visualization of the physical query plan. The physical query plan in Myria is composed of
-a DAG of fragments, where each DAG link represents communication between workers, possibly over the network. 
-\system displays this DAG in an interactive fashion with clickable elements for fragment nodes and communication links. 
-This DAG also represents a high-level data flow in the DDBMS.
+\item An interactive visualization of the physical query plan. The physical
+query plan is composed of a DAG of query \emph{fragments}. A query
+fragment is a subtree of the query execution plan that is separated from the
+rest of the plan by a network boundary. This means that the operators inside a
+query fragment can be executed by a single thread. Each fragment is scheduled
+to run on multiple workers, which means each worker thread executes the same set of
+operators inside the fragment, but on different data. Each DAG link represents
+communication between workers, possibly over the network.  \system displays
+this DAG in an interactive fashion with clickable elements for fragment nodes
+and communication links.  This DAG also represents a high-level data flow in
+the DDBMS.
 
-\item A view for cluster utilization and operator performance for each \emph{Myria} segment. Each segment is composed of
-multiple sequential database operators running on the same worker. Upon clicking on a collapsed segment node in the
-query plan DAG, the node expands to display the operators comprising the segment and renders the fragment execution view.
+\item A view for cluster utilization and operator performance for each query fragment. Each fragment is composed of
+multiple sequential database operators running on the same worker. Upon clicking on a collapsed fragment node in the
+query plan DAG, the node expands to display the operators comprising the fragment and renders the fragment execution view.
 The fragment execution view visualizes the utilization of the cluster by showing whether each worker is busy working 

### 2014-03-19 final-report/paper.tex
network vis section: still work to do

--- 2014-03-19 final-report/paper.tex
@@
 
-% Umar
+The Worker Communication view consists of two charts: the aggregate communication 
+chart and the time series chart. The aggregate communication chart is a matrix representing
+a cluster of $N$ workers. Each cell $m,n$ in the matrix represents the total number of tuples sent
+by worker $m$ to worker $n$ during the query execution. Figure \ref{fig:matrix} shows this chart along 
+with the selected link in the query plan DAG.
 
-\todo{write me}
+\begin{figure}[ht]
+  \includegraphics[width=\columnwidth]{images/networkVis1}
+  \caption{Worker communication matrix chart.}
+  \label{fig:matrix}
+\end{figure}
 

### 2014-03-20 final-report/paper.tex
Fixes to future work & other nits

--- 2014-03-19 final-report/paper.tex
@@
 database programmer or administrator could quickly discover run-time bottlenecks as well as problematic data skew issues, 
-saving them significant guesswork and programming effort to identify the appropriate logs to analyze. Our visualization tool
-is called \system, which uses profiling logs from \emph{Myria} \cite{myria} as the prototype DDMS. Myria is a distributed 
+saving them significant guesswork and programming effort to identify the appropriate logs to analyze. Our visualization tool,
+\system, uses profiling logs from \emph{Myria} \cite{myria} as the prototype DDMS. Myria is a distributed 
 database platform that provides \emph{big data management and analytics as a service} primarily for scientific applications.
@@
 
+This paper introduced \system, a visualization tool carefully tailored to help
+database developers or users to quickly identify common distributed databases
+performance bottlenecks, like tail latencies, data skew, etc. Our tool has been
+successfully integrated with the Myria distributed database system, and has already
+been used to identify several bugs, like poor physical query plan optimization,
+tail latencies, poor data partitioning. We believe the visualization techniques
+we used are appropriate to help the user quickly navigate to/find the cause of the
+performance problem. \system could be easily plugged-in with another
+distributed database system, the only requirement being a specific log data
+format.
+
 % Capabilities and Limitations
-
-\todo{write me}
 
@@
 
-A focus of the current implementation is on scalability to many workers and long running queries.
-More aggregations have to be done by the database itself.
-Most importantly, in the fragment visualization that shows spans for all events we should limit the data that is being downloaded and rendered to what the user can see.
-With the architecture as explained in Section~\ref{sec:back}, we are able to express these filter conditions as queries.
+The current implementation focuses on scalability. We managed to some extent to
+address the scalability issues raised by big query plans, large amount of
+workers or long running queries. For example the \network view uses a
+scalable heatmap to show the communication between all the workers in the
+system. However, other techniques have to be further implemented to address
+this issue. One example would be a more interactive query plan graph, which the
+users can zoom. Another example, in the \fragment view that shows spans
+for all events we should limit the data that is being downloaded and rendered
+to what the user can see.
 
-Besides performance improvements, we plan to integrate X-trace and its visualization to offer the orthogonal view and allow users to trace how tuple batches flow through the operators and between operators.
-Combined with the visualizations described in this paper, it will form a powerful debugging tool that will help us to better understand and improve query execution.
+Besides performance improvements, we plan to integrate X-trace and its
+visualization to offer the orthogonal view that allows users to trace how tuple
+batches flow through the operators and between operators. Combined with the
+visualizations described in this paper, it will form a powerful debugging tool
+that will help uers better understand and improve query execution.
 

### 2014-03-20 final-report/paper.tex
network viz section

--- 2014-03-19 final-report/paper.tex
@@
 by worker $m$ to worker $n$ during the query execution. Figure \ref{fig:matrix} shows this chart along 
-with the selected link in the query plan DAG.
+with the selected link in the query plan DAG. Hovering the mouse over a cell brings up a tooltip
+stating the total number of tuples exchanged between the source and destination workers. The intensity
+of the cell color reflects the number of tuples exchanged, i.e., a larger number translates into a darker color 
+and a smaller number results in a lighter color. The color pallet used is immune to color-blindness issues.
+The matrix chart representing aggregate communication 
+between workers help uncover data skew issues, as discussed in \S \ref{sec:eval}.
+
+We initially considered having a chord diagram\footnote{\url{http://bl.ocks.org/mbostock/4062006}} instead of a matrix 
+for the aggregate communication chart. But feedback during the progress presentation session convinced us that 
+a chord diagram would suffer from scalability problems. Adding an option for displaying a chord diagram remains
+part of future work.
 
@@
   \includegraphics[width=\columnwidth]{images/networkVis1}
-  \caption{Worker communication matrix chart.}
+  \caption{Worker communication matrix chart. Each cell represents the number of exchanged tuples between a 
+  pair of workers.}
   \label{fig:matrix}
@@
 \section{Evaluation}
+\label{sec:eval}
 

### 2014-03-20 final-report/paper.tex
networkvis 2
rest of networkviz section

@@
 of the cell color reflects the number of tuples exchanged, i.e., a larger number translates into a darker color 
-and a smaller number results in a lighter color. The color pallet used is immune to color-blindness issues.
+and a smaller number results in a lighter color. The color pallet used is grayscale and hence
+immune to color-blindness issues. The color scale uses LAB interpolation.
 The matrix chart representing aggregate communication 
@@
 
-\begin{figure}[ht]
+\begin{figure}[!ht]
   \includegraphics[width=\columnwidth]{images/networkVis1}
@@
   \label{fig:matrix}
+\end{figure}
+
+When the user clicks on a matrix cell, they are able to view the breakdown of the worker pair communication in a 
+time series chart below the matrix. The user can select multiple such cells to compare different time 
+series. Figure \ref{fig:ts} shows an example when the user selected two pairs: (8-2) and (7-6). The x-axis shows
+time in seconds and y-axis shows number of tuples exchanged between the source and destination workers.
+Just looking at the matrix chart it is evident that there is more data exchanged between the source-destination pair
+(8-2) than the pair (7-6). But as the time series comparison shows, there is more communication between (7-6)
+than (8-2) for the last 2.5 seconds of the query execution. This points to interesting patterns as to the rate at which 
+work is done at different workers at different stages of the execution. Distinct points on the lines are highlighted as
+black dots. Hovering the cursor over a dot shows the information for that data point in a tooltip. Meanwhile the cells 
+in the matrix remain highlighted with the same color used to draw the corresponding time series line. This aids in the
+user keeping track of the mapping between the matrix cell and the corresponding time series line. We also provide 
+a 'clear selection' button that erases everything from the time series chart upon clicking. Finally we provide the
+functionality whereby clicking a row or column label selects every cell in that row or column and displays the 
+corresponding time series in the time series chart.
+
+\begin{figure}[!ht]
+  \includegraphics[width=\columnwidth]{images/networkVis2}
+  \caption{Worker communication time series chart.}
+  \label{fig:ts}
 \end{figure}

### 2014-03-20 final-report/paper.tex
network vis 2 caption
Thierry and I fixed numbering scheme. Other small fixes

--- 2014-03-20 final-report/paper.tex
@@
 been used to identify several bugs, like poor physical query plan optimization,
-tail latencies, poor data partitioning. We believe the visualization techniques
-we used are appropriate to help the user quickly navigate to/find the cause of the
-performance problem. \system could be easily plugged-in with another
-distributed database system, the only requirement being a specific log data
-format.
+tail latencies, poor data partitioning, which led us to believe the
+visualization techniques we used are appropriate to help the user quickly
+navigate to/find the cause of the performance problem. \system could be easily
+plugged-in with another distributed database system, the only requirement being
+a specific log data format.
 

### 2014-03-20 final-report/paper.tex

--- 2014-03-20 final-report/paper.tex
@@
   \includegraphics[width=\columnwidth]{images/networkVis2}
-  \caption{Worker communication time series chart.}
+  \caption{Worker communication time series chart. The user can select multiple cells in the matrix to compare the
+  corresponding time series.}
   \label{fig:ts}

### 2014-03-20 final-report/paper.tex
Figure instead of Fig.

@@
 The \fragment view comprises two charts: the utilization chart and the
-operators chart. The utilization chart (Fig.~\ref{fig:utilization_chart}) shows on
+operators chart. The utilization chart (Figure~\ref{fig:utilization_chart}) shows on
 how many workers a fragment is running over time. The chart is used to
@@
 analyze the query execution by displaying the schedule of each of the worker
-nodes, in the operators chart (Fig.~\ref{fig:operators_chart}). The per-worker
+nodes, in the operators chart (Figure~\ref{fig:operators_chart}). The per-worker
 schedule shows what operators the worker executed and for how long.
@@
 a cluster of $N$ workers. Each cell $m,n$ in the matrix represents the total number of tuples sent
-by worker $m$ to worker $n$ during the query execution. Figure \ref{fig:matrix} shows this chart along 
+by worker $m$ to worker $n$ during the query execution. Figure~\ref{fig:matrix} shows this chart along 
 with the selected link in the query plan DAG. Hovering the mouse over a cell brings up a tooltip
@@
 time series chart below the matrix. The user can select multiple such cells to compare different time 
-series. Figure \ref{fig:ts} shows an example when the user selected two pairs: (8-2) and (7-6). The x-axis shows
+series. Figure~\ref{fig:ts} shows an example when the user selected two pairs: (8-2) and (7-6). The x-axis shows
 time in seconds and y-axis shows number of tuples exchanged between the source and destination workers.

### 2014-03-20 final-report/paper.tex
Typos and new section

--- 2014-03-20 final-report/paper.tex
@@
 \begin{itemize}
+
+\todo{it's not a dag, as we noticed}
 
@@
 quickly reveal any patterns in the schedule, like tail latency, long periods of
-idleness, etc. We implemented a small brush (at the top of the figure) that
+idleness, etc. Execution skew is visible as humps.
+We implemented a small brush (at the top of the figure) that
 allows the user to zoom-in and further analyze the problematic area.
@@
 The \overall shows small multiples of the utilization chart as described in the previous section.
-Execution skew can be spotted in this view. Furthermore, the user can compare the fragments and find
+In addition to identifying execution skew, the user can compare the fragments and find
 correlations between the execution of different fragments. This view helps the user navigate to the
@@
 
+\todo{is this figure necessary or is the figure in the evaluation section enough?}
+
 \begin{figure}[ht]
@@
 This query joins Twitter followers and followees by user ID.
-The query can be written in SQL as expressed below.
+The query can be written in SQL as:
 
@@
 
-%dom
+The \overall allows the user to quickly get an overview over the execution of all fragments
+on all workers in the form of small multiples of the utilization charts as introduced in
+Section~\ref{sec:fragment}. In Figure~\ref{fig:overview_skew} the user can see for example
+that fragment~2 is very irregularly executed on different workers. If the graph is not a
+single hill (as it is the case in Fragment 1, 3 and 4), workers are sleeping because they
+either don't have enough data available to execute the operators or the output buffers
+are full and data cannot be sent to the next fragment in the query graph.
+
+The user has not identifies Fragment~2 to be a performance bottleneck and can use the other views
+to further investigate its causes.
+
+\begin{figure}[ht]
+  \includegraphics[width=\columnwidth]{images/overview_skew}
+  \caption{This \overall shows very irregular usage in fragment 2.}
+  \label{fig:overview_skew}
+\end{figure}
 
@@
 
+\todo{Let's not call the fragments frag2 but write fragment 2}
+
 The \network visualization at a glance shows how much data was sent between workers by one query fragment to the next.
-The two \network views in Figure~\ref{fig:no-skew} and Figure~\ref{fig:skew} were obtained by selecting the Frag0$\rightarrow$Frag2 and Frag1$\rightarrow$Frag2 edges respectively in the \graph view.
+The two \network views in Figure~\ref{fig:network} were obtained by selecting the Frag0$\rightarrow$Frag2 and Frag1$\rightarrow$Frag2 edges respectively in the \graph view.
 Figure~\ref{fig:network} displays aggregate data communication between workers.
-In Figure~\ref{fig:no-skew} the amount of data sent from each worker to every other worker is balanced as opposed to Figure~\ref{fig:skew} where one can see that the volume of data sent to worker~1 and worker~5 outweights data sent to the other workers in the system.
+In Figure~\ref{fig:no-skew} the amount of data sent from each worker to every other worker is balanced as opposed to Figure~\ref{fig:skew} where one can see that the volume of data sent to worker~1 and worker~5 outweighs data sent to the other workers in the system.
 This apparent data skew could be caused by partitioning issues.
@@
 The color scheme used for each worker schedules in the \fragment view is color-coded to match the operators in Figure~\ref{fig:join}.
-Consequently, one can quickly observe in the \fragment view that worker~5 is busy sending join results from t=2.5s to t=5.0s, which differes from the rest of the workers.
+Consequently, one can quickly observe in the \fragment view that worker~5 is busy sending join results from $t = 2.5s$ to $t = 5.0s$, which differs from the rest of the workers.
 Using similar reasoning, one can conclude that worker 5 presents a performance bottleneck in the execution of the query on Myria.
-Perhaps the user can modify the way the input data set gets partitioned.
+In the short term, the user can modify the way the input data set gets partitioned.
 Alternatively she could change the number of workers allocated for the job.
+in the long term, developers should look into ways to improve data flow to avoid that workers have to wait for input data to process.
 
@@
 text-based performance logs, and runtimes.
-Being able to navigate performance data using a simple interface can indeed facilitate the task of a developper who is designing Myria's back end, or a Myria user who is writing complex queries to process large amounts of data.
+Being able to navigate performance data using a simple interface can indeed facilitate the task of a developer who is designing Myria's back end, or a Myria user who is writing complex queries to process large amounts of data.
 That said, \system can only help the user visualize query execution performance, but won't necessarily point the user to what she needs to do to improve performance of her query execution.
@@
 been used to identify several bugs, like poor physical query plan optimization,
-tail latencies, poor data partitioning, which led us to believe the
-visualization techniques we used are appropriate to help the user quickly
-navigate to/find the cause of the performance problem. \system could be easily
-plugged-in with another distributed database system, the only requirement being
-a specific log data format.
+tail latencies, poor data partitioning. We believe the visualization techniques
+we used are appropriate to help the user quickly navigate to/find the cause of the
+performance problem. \system could be easily plugged-in with another
+distributed database system, the only requirement being a specific log data
+format.
 

### 2014-03-20 final-report/paper.tex

--- 2014-03-20 final-report/paper.tex
@@
 part of future work.
+
+\todo{is this figure necessary or is the figure in the evaluation section enough?}
 

### 2014-03-20 final-report/paper.tex
Sorry
Ref and typo

@@
 
+\todo{it's not a dag, as we noticed}
+
 \begin{itemize}
-
-\todo{it's not a dag, as we noticed}
 
@@
 
-\textbf{Graph Rendering:} We use D3 \cite{d3} to render the \graph graph, which allows us to support various interactions and transistions.
+\textbf{Graph Rendering:} We use D3 \cite{2011-d3} to render the \graph graph, which allows us to support various interactions and transitions.
 We use GraphViz \cite{Ellson01graphviz} in the backend to generate graph layout information.
@@
 Expanding or reducing a node can cause large changes in the graph layout as GraphViz changes the graph layout to minimize overall area.
-This graph reshuffeling is illustrated in Figure~\ref{graph} where expanding Frag2 causes the layout to change.
+This graph re-shuffling is illustrated in Figure~\ref{graph} where expanding Frag2 causes the layout to change.
 To address these layout changes, we implemented transitions to allow the user to track the fragments as those get reshuffled.
@@
 
-% use the word spans when referring to events with duration (used in dapper)
+\todo{use the word spans when referring to events with duration (used in dapper)}
 

### 2014-03-20 final-report/paper.tex
Rewrote abstract

--- 2014-03-20 final-report/paper.tex
@@
 
-We propose a visualization system to explore and understand query
-execution and data movement in a distributed database management system
-(DDBMS). Our tool provides insight into: (1) data flow between query segments
-and between workers, (2) query execution and operator dependencies, (3) cluster
-utilization, (4) network utilization. Our tool, \system, helps
-developers understand and improve query execution, and provides insight into
-common problems such as data skew or performance bottlenecks.
+We propose \system, a visualization tool that helps database developers/users
+explore and understand query execution and data movement in a distributed
+database management system (DDBMS). \system provides quick insight into common
+problems such as data skew or performance bottleneck by leveraging
+visualization techniques to present: (1) data flow between query segments and
+between workers, (2) query execution and operator dependencies, (3) cluster
+utilization, (4) network utilization.
 
@@
 of query execution.
+\todo{it's not a dag, as we noticed}
 
 \begin{itemize}
-
-\todo{it's not a dag, as we noticed}
 

### 2014-03-20 final-report/paper.tex
reworked some parts

--- 2014-03-20 final-report/paper.tex
@@
 
-\todo{it's not a dag, as we noticed}
-
 \begin{itemize}
@@
 
-\system's architecture (Figure~\ref{fig:arch}) consists of a back-end,
-used as a plug-in interface for the distributed database system targeted, and a
-front-end which implements the views described above. The front-end and
+\system's architecture (Figure~\ref{fig:arch}) is composed of: (1) a back-end
+used as a plug-in interface for the targeted distributed database system, and (2) a
+front-end that produces performance visualizations what we describe in Section~\ref{sec:front}. The front-end and
 back-end of \system are highly decoupled so that the visualization can be used not only with
@@
 
-The task of the back-end is to log when operators are called, when the call returns and when data is sent to another worker.
-Using this data and simple transformations, we can obtain the information necessary to create the visualizations to answer the questions in Section~\ref{sec:approach}.
+The role of the back-end is produce event logs that record when operators are called, when the call returns and when data is sent to another worker.
+These logs are then used to create the visualizations in \system's web UI as described in Section~\ref{sec:front}.
 
-There are two possible approaches that we explored to collect event logs that can be used to extract spans and other performance related data necessary for the visualization.
-First, events could be logged to files using a standard logging tool and then collect the files on one node using remote copy.
-Once the logs are on one machine, the logs can be parsed and the data can be transformed into the desired format.
+There are two possible approaches that we explored to collect event logs.
+The first and simplest approach was to write the logs on files directly to disk.
+The files would then get hauled back on one node using remote copy.
+The logs would then be parsed, and the data transformed into the desired format.
 
-We switched to a different approach as shown in Figure~\ref{fig:arch} where logs are written directly into the worker's database.
-When the front-end requests the data, the master executes a query that collects the data and streams the results directly to the requesting client.
-Logs are thus tuples in a relation and no longer separate in files and in a custom format.
-Despite being more reliable, faster and using the same logic that Myria already implements, we are able to write queries to aggregate, filter and transform the collected data.
+We switched to a more scalable data collection approach approach as shown in Figure~\ref{fig:arch} where logs get written directly into the worker's database as tuples in a relation.
+When the front-end client requests data, the master executes a query that collects the data and streams the results directly to the requesting client.
+We found this data collection approach to be more scalable, reliable and faster.
+More particularly, we are able to write queries to aggregate, filter and transform the collected data.
 
-For example, the data for how many workers are working on a certain fragment as described in Section~\ref{sec:fragments} is generated from the event log using a query to select only events for the root operators (the operators that have no parent in the same fragment) in each fragment and a custom map function that carries state.
+For example, when querying how many workers are working on a certain fragment at a given time to produce the visualization described in Section~\ref{sec:fragments}, we used a query to select only events in the root operators (the operators that have no parent in the same fragment) for each fragment and a custom map function that carries state.
 The state is an integer that is incremented when the operator is called on a worker and decremented when the operator returns.
@@
 \subsection{Front-end}
+\label{sec:front}
 
@@
 D3 is a JavaScript framework for data visualization on the web.
-We used visualization techniques such as \emph{focus+context}\cite{furnas1986generalized} that possible to implement conveniently using D3.
+We used visualization techniques such as \emph{focus+context}\cite{furnas1986generalized} that were made possible using D3.
 
-\system's front-end gets data from the Myria server as JSON and CSV files.
+\system's front-end gets data from the Myria server using \system's back-end as JSON and CSV files.
 JSON objects are used to describe the physical queryplan.
@@
 
-\textbf{Graph Rendering:} We use D3 \cite{2011-d3} to render the \graph graph, which allows us to support various interactions and transitions.
+\textbf{Graph Rendering:} We use D3 \cite{d3} to render the \graph graph, which allows us to support various interactions and transistions.
 We use GraphViz \cite{Ellson01graphviz} in the backend to generate graph layout information.
@@
 Expanding or reducing a node can cause large changes in the graph layout as GraphViz changes the graph layout to minimize overall area.
-This graph re-shuffling is illustrated in Figure~\ref{graph} where expanding Frag2 causes the layout to change.
+This graph reshuffeling is illustrated in Figure~\ref{graph} where expanding Frag2 causes the layout to change.
 To address these layout changes, we implemented transitions to allow the user to track the fragments as those get reshuffled.
@@
 quickly reveal any patterns in the schedule, like tail latency, long periods of
-idleness, etc. Execution skew is visible as humps.
-We implemented a small brush (at the top of the figure) that
+idleness, etc. We implemented a small brush (at the top of the figure) that
 allows the user to zoom-in and further analyze the problematic area.
@@
 
-\todo{use the word spans when referring to events with duration (used in dapper)}
+% use the word spans when referring to events with duration (used in dapper)
 
@@
 The \overall shows small multiples of the utilization chart as described in the previous section.
-In addition to identifying execution skew, the user can compare the fragments and find
+Execution skew can be spotted in this view. Furthermore, the user can compare the fragments and find
 correlations between the execution of different fragments. This view helps the user navigate to the
@@
 part of future work.
-
-\todo{is this figure necessary or is the figure in the evaluation section enough?}
 
@@
 This query joins Twitter followers and followees by user ID.
-The query can be written in SQL as:
+The query can be written in SQL as expressed below.
 
@@
 
-The \overall allows the user to quickly get an overview over the execution of all fragments
-on all workers in the form of small multiples of the utilization charts as introduced in
-Section~\ref{sec:fragment}. In Figure~\ref{fig:overview_skew} the user can see for example
-that fragment~2 is very irregularly executed on different workers. If the graph is not a
-single hill (as it is the case in Fragment 1, 3 and 4), workers are sleeping because they
-either don't have enough data available to execute the operators or the output buffers
-are full and data cannot be sent to the next fragment in the query graph.
-
-The user has not identifies Fragment~2 to be a performance bottleneck and can use the other views
-to further investigate its causes.
-
-\begin{figure}[ht]
-  \includegraphics[width=\columnwidth]{images/overview_skew}
-  \caption{This \overall shows very irregular usage in fragment 2.}
-  \label{fig:overview_skew}
-\end{figure}
+%dom
 
@@
 
-\todo{Let's not call the fragments frag2 but write fragment 2}
-
 The \network visualization at a glance shows how much data was sent between workers by one query fragment to the next.
-The two \network views in Figure~\ref{fig:network} were obtained by selecting the Frag0$\rightarrow$Frag2 and Frag1$\rightarrow$Frag2 edges respectively in the \graph view.
+The two \network views in Figure~\ref{fig:no-skew} and Figure~\ref{fig:skew} were obtained by selecting the Frag0$\rightarrow$Frag2 and Frag1$\rightarrow$Frag2 edges respectively in the \graph view.
 Figure~\ref{fig:network} displays aggregate data communication between workers.
-In Figure~\ref{fig:no-skew} the amount of data sent from each worker to every other worker is balanced as opposed to Figure~\ref{fig:skew} where one can see that the volume of data sent to worker~1 and worker~5 outweighs data sent to the other workers in the system.
+In Figure~\ref{fig:no-skew} the amount of data sent from each worker to every other worker is balanced as opposed to Figure~\ref{fig:skew} where one can see that the volume of data sent to worker~1 and worker~5 outweights data sent to the other workers in the system.
 This apparent data skew could be caused by partitioning issues.
@@
 The color scheme used for each worker schedules in the \fragment view is color-coded to match the operators in Figure~\ref{fig:join}.
-Consequently, one can quickly observe in the \fragment view that worker~5 is busy sending join results from $t = 2.5s$ to $t = 5.0s$, which differs from the rest of the workers.
+Consequently, one can quickly observe in the \fragment view that worker~5 is busy sending join results from t=2.5s to t=5.0s, which differes from the rest of the workers.
 Using similar reasoning, one can conclude that worker 5 presents a performance bottleneck in the execution of the query on Myria.
-In the short term, the user can modify the way the input data set gets partitioned.
+Perhaps the user can modify the way the input data set gets partitioned.
 Alternatively she could change the number of workers allocated for the job.
-in the long term, developers should look into ways to improve data flow to avoid that workers have to wait for input data to process.
 
@@
 text-based performance logs, and runtimes.
-Being able to navigate performance data using a simple interface can indeed facilitate the task of a developer who is designing Myria's back end, or a Myria user who is writing complex queries to process large amounts of data.
+Being able to navigate performance data using a simple interface can indeed facilitate the task of a developper who is designing Myria's back end, or a Myria user who is writing complex queries to process large amounts of data.
 That said, \system can only help the user visualize query execution performance, but won't necessarily point the user to what she needs to do to improve performance of her query execution.

### 2014-03-20 final-report/paper.tex
Move some content

@@
 
-For example, when querying how many workers are working on a certain fragment at a given time to produce the visualization described in Section~\ref{sec:fragments}, we used a query to select only events in the root operators (the operators that have no parent in the same fragment) for each fragment and a custom map function that carries state.
+For example, when querying how many workers are working on a certain fragment at a given time to produce the visualization described in Section~\ref{sec:fragment}, we used a query to select only events in the root operators (the operators that have no parent in the same fragment) for each fragment and a custom map function that carries state.
 The state is an integer that is incremented when the operator is called on a worker and decremented when the operator returns.
@@
 
-We highly suggest that developers use the second approach when porting \system to a new DDBMS.
-
 \subsection{Front-end}
@@
 D3 is a JavaScript framework for data visualization on the web.
-We used visualization techniques such as \emph{focus+context}\cite{furnas1986generalized} that were made possible using D3.
 
-\system's front-end gets data from the Myria server using \system's back-end as JSON and CSV files.
-JSON objects are used to describe the physical queryplan.
+\system's front-end gets data from the Myria server as JSON and CSV files.
+JSON is used to describe the physical queryplan.
 CSV files contain logs on query execution at the fragment and operator level as well as data exchange between workers at different stages of the query execution.
@@
 idleness, etc. Execution skew is visible as humps.
-We implemented a small brush (at the top of the figure) that
-allows the user to zoom-in and further analyze the problematic area.
+We implemented \emph{focus+context}\cite{furnas1986generalized} in form of
+a small brush (at the top of the figure) that allows the user to zoom-in and further
+analyze the problematic area without loosing the context of the whole execution.
 
@@
   \includegraphics[width=\columnwidth]{images/utilization_chart}
-  \caption{The utilization chart. }
+  \caption{The utilization chart.}
   \label{fig:utilization_chart}
@@
 \subsubsection{Overview over all fragments}
-\label{sec:fragments}
+\label{sec:overview}
 
@@
 plugged-in with another distributed database system, the only requirement being
-a specific log data format.
+a specific log data format. We explained how we collect logs in Myria and recommend
+our approach to developers that are porting \system to a new DDBMS.
 

### 2014-03-20 final-report/paper.tex
Tackled the background

--- 2014-03-20 final-report/paper.tex
@@
 In particular, \system is built to inspect query execution in
-Myria\footnote{\url{http://myria.cs.washington.edu/}}, a distributed big data
+Myria~\cite{myria}, a distributed big data
 management system currently being developed in the UW CSE database group. Myria
@@
 saving them significant guesswork and programming effort to identify the appropriate logs to analyze. Our visualization tool,
-\system, uses profiling logs from \emph{Myria} \cite{myria} as the prototype DDMS. Myria is a distributed
+\system, uses profiling logs from Myria as the prototype DDMS. Myria is a distributed
 database platform that provides \emph{big data management and analytics as a service} primarily for scientific applications.
@@
 
-Twitter developed Ambrose\cite{ambrose}, a platform for visualization and real-time monitoring of MapReduce data workflows.
-They offer three different views to show associated jobs, job dependencies and progress.
-Ambrose has been released as open source.
-Ambrose, however, is not suitable for our needs as the abstraction level of jobs is too high and does not capture single operators.
+Ambrose\cite{ambrose} is a platform developed by Twitter to visualize and
+real-time monitor MapReduce data workflows. Ambrose
+offers three different views to show associated jobs, job dependencies and
+progress, which are not suitable for our needs as the abstraction level of jobs
+is too high and does not capture single operators.
 
-Google's Dapper\cite{sigelman2010dapper} is a distributed systems tracing infrastructure and offers fine grained tracing of calls in Google's distributed systems.
-They also proposed an interface for visualizing traces.
-Similarly, X-trace\cite{fonseca2007x} was developed as a framework to trace which events cause what other events in a distributed environment.
-recently, there has been work on visualizing event traces collected in X-trace\footnote{\url{https://github.com/brownsys/X-Trace/tree/master/src/webui/html/interactive}}.
-Due to the importance of these kinds of debug facilities, Twitter closely modeled Zipkin\cite{zipkin} after Dapper and X-trace and released it as open source.
+Google's Dapper\cite{sigelman2010dapper} is a distributed systems tracing
+infrastructure that offers fine-grained tracing of calls in Google's distributed
+systems. They also proposed an interface for visualizing traces. Similarly,
+X-trace\cite{fonseca2007x} was developed as a framework to trace which events
+cause what other events in a distributed environment. Recently, there has been
+work on visualizing event traces collected in
+X-trace\footnote{\url{https://github.com/brownsys/X-Trace/tree/master/src/webui/html/interactive}}.
+Due to the importance of these kinds of debug facilities, Twitter closely
+modeled Zipkin\cite{zipkin} after Dapper and X-trace and released it as open
+source.
 
-Dapper and X-trace focus on how data flows through a distributed system.
-In \system the visualization focuses on the operators and how the data flows through them.
-This orthogonal view is better suited for debugging performance bottlenecks in DDBMS and also scales to a larger number of events.
-Furthermore, in contrast to  we offer different abstraction levels, which enables users to find problems faster and handle larger amounts of profiling data.
-Also, \system is specifically designed to help developers understand query execution in distributed database system like Myria as opposed to general traces in distributed systems.
+Dapper and X-trace focus on how data flows through a distributed system. In
+contrast, in \system the visualization focuses on the operators and how the
+data flows through them. This orthogonal view is better suited for debugging
+performance bottlenecks in DDBMSs and also scales to a larger number of events.
+Furthermore, \system offers different abstraction levels, which enables users
+to find problems faster and handle larger amounts of profiling data. Also,
+\system is specifically designed to help developers understand query execution
+in distributed database system as opposed to general traces in distributed
+systems.
 
-Tools to visualize query plans for example in SQL Server, as used to improve performance for the SDSS Sky survey\cite{szalay2002sdss}, focus on optimizing queries and not query execution and have no visualization of data flow, which is necessary to optimize physical query execution.
+Tools to visualize query plans, for example those used to improve the
+performance for the SDSS Sky survey\cite{szalay2002sdss}, focus on optimizing
+queries and not query execution and have no visualization of data flow, which
+is necessary to optimize physical query execution.
 
@@
 
-\system's architecture (Figure~\ref{fig:arch}) is composed of: (1) a back-end
-used as a plug-in interface for the targeted distributed database system, and (2) a
-front-end that produces performance visualizations what we describe in Section~\ref{sec:front}. The front-end and
+\system's architecture (Figure~\ref{fig:arch}) consists of a back-end,
+used as a plug-in interface for the distributed database system targeted, and a
+front-end which implements the views described above. The front-end and
 back-end of \system are highly decoupled so that the visualization can be used not only with
@@
 
-The role of the back-end is produce event logs that record when operators are called, when the call returns and when data is sent to another worker.
-These logs are then used to create the visualizations in \system's web UI as described in Section~\ref{sec:front}.
+The task of the back-end is to log when operators are called, when the call returns and when data is sent to another worker.
+Using this data and simple transformations, we can obtain the information necessary to create the visualizations to answer the questions in Section~\ref{sec:approach}.
 
-There are two possible approaches that we explored to collect event logs.
-The first and simplest approach was to write the logs on files directly to disk.
-The files would then get hauled back on one node using remote copy.
-The logs would then be parsed, and the data transformed into the desired format.
+There are two possible approaches that we explored to collect event logs that can be used to extract spans and other performance related data necessary for the visualization.
+First, events could be logged to files using a standard logging tool and then collect the files on one node using remote copy.
+Once the logs are on one machine, the logs can be parsed and the data can be transformed into the desired format.
 
-We switched to a more scalable data collection approach approach as shown in Figure~\ref{fig:arch} where logs get written directly into the worker's database as tuples in a relation.
-When the front-end client requests data, the master executes a query that collects the data and streams the results directly to the requesting client.
-We found this data collection approach to be more scalable, reliable and faster.
-More particularly, we are able to write queries to aggregate, filter and transform the collected data.
+We switched to a different approach as shown in Figure~\ref{fig:arch} where logs are written directly into the worker's database.
+When the front-end requests the data, the master executes a query that collects the data and streams the results directly to the requesting client.
+Logs are thus tuples in a relation and no longer separate in files and in a custom format.
+Despite being more reliable, faster and using the same logic that Myria already implements, we are able to write queries to aggregate, filter and transform the collected data.
 
-For example, when querying how many workers are working on a certain fragment at a given time to produce the visualization described in Section~\ref{sec:fragments}, we used a query to select only events in the root operators (the operators that have no parent in the same fragment) for each fragment and a custom map function that carries state.
+For example, the data for how many workers are working on a certain fragment as described in Section~\ref{sec:fragments} is generated from the event log using a query to select only events for the root operators (the operators that have no parent in the same fragment) in each fragment and a custom map function that carries state.
 The state is an integer that is incremented when the operator is called on a worker and decremented when the operator returns.
@@
 \subsection{Front-end}
-\label{sec:front}
 
@@
 D3 is a JavaScript framework for data visualization on the web.
-We used visualization techniques such as \emph{focus+context}\cite{furnas1986generalized} that were made possible using D3.
+We used visualization techniques such as \emph{focus+context}\cite{furnas1986generalized} that possible to implement conveniently using D3.
 
-\system's front-end gets data from the Myria server using \system's back-end as JSON and CSV files.
+\system's front-end gets data from the Myria server as JSON and CSV files.
 JSON objects are used to describe the physical queryplan.

### 2014-03-20 final-report/paper.tex
Cool title
edits on case 1

--- 2014-03-20 final-report/paper.tex
@@
 
-For example, when querying how many workers are working on a certain fragment at a given time to produce the visualization described in Section~\ref{sec:fragment}, we used a query to select only events in the root operators (the operators that have no parent in the same fragment) for each fragment and a custom map function that carries state.
+For example, when querying how many workers are working on a certain fragment at a given time to produce the visualization described in Section~\ref{sec:fragments}, we used a query to select only events in the root operators (the operators that have no parent in the same fragment) for each fragment and a custom map function that carries state.
 The state is an integer that is incremented when the operator is called on a worker and decremented when the operator returns.
@@
 
+We highly suggest that developers use the second approach when porting \system to a new DDBMS.
+
 \subsection{Front-end}
@@
 D3 is a JavaScript framework for data visualization on the web.
+We used visualization techniques such as \emph{focus+context}\cite{furnas1986generalized} that were made possible using D3.
 
-\system's front-end gets data from the Myria server as JSON and CSV files.
-JSON is used to describe the physical queryplan.
+\system's front-end gets data from the Myria server using \system's back-end as JSON and CSV files.
+JSON objects are used to describe the physical queryplan.
 CSV files contain logs on query execution at the fragment and operator level as well as data exchange between workers at different stages of the query execution.
@@
 idleness, etc. Execution skew is visible as humps.
-We implemented \emph{focus+context}\cite{furnas1986generalized} in form of
-a small brush (at the top of the figure) that allows the user to zoom-in and further
-analyze the problematic area without loosing the context of the whole execution.
+We implemented a small brush (at the top of the figure) that
+allows the user to zoom-in and further analyze the problematic area.
 
@@
   \includegraphics[width=\columnwidth]{images/utilization_chart}
-  \caption{The utilization chart.}
+  \caption{The utilization chart. }
   \label{fig:utilization_chart}
@@
 \subsubsection{Overview over all fragments}
-\label{sec:overview}
+\label{sec:fragments}
 
@@
 
-The \overall shows small multiples of the utilization chart as described in the previous section.
+The \overall shows small multiples of the utilization chart described in the previous section.
 In addition to identifying execution skew, the user can compare the fragments and find
-correlations between the execution of different fragments. This view helps the user navigate to the
-right fragment and is not intended to offer further inside into the causes or performance problems.
+correlations between the execution of different fragments. This view helps the user navigate between 
+fragments and is not intended to offer deep insight into performance issues. Section~{\sec:case1} describes
+a usecase for the \overall visualization.
 
@@
 \subsection{Case 1: Identifying Performance Bottlenecks using the \overall}
+\label{sec:case1}
 
-The \overall allows the user to quickly get an overview over the execution of all fragments
-on all workers in the form of small multiples of the utilization charts as introduced in
-Section~\ref{sec:fragment}. In Figure~\ref{fig:overview_skew} the user can see for example
-that fragment~2 is very irregularly executed on different workers. If the graph is not a
-single hill (as it is the case in Fragment 1, 3 and 4), workers are sleeping because they
-either don't have enough data available to execute the operators or the output buffers
-are full and data cannot be sent to the next fragment in the query graph.
-
-The user has not identifies Fragment~2 to be a performance bottleneck and can use the other views
-to further investigate its causes.
+The \overall view provides the user with a summary of the execution of all fragments
+on all workers as described in
+Section~\ref{sec:fragment}. In Figure~\ref{fig:overview_skew} the user can see for instance
+that Fragment~2 is irregularly executed on multiple workers. 
+If the graph is not monolithic (as it is the case in Fragment 1, 3 and 4), workers are sleeping because they
+either (1) don't have enough data available to work on (data dependency problem) or (2) the destination worker's
+input buffers are full thus forcing the source workers to stall.
 
@@
 plugged-in with another distributed database system, the only requirement being
-a specific log data format. We explained how we collect logs in Myria and recommend
-our approach to developers that are porting \system to a new DDBMS.
+a specific log data format.
 

### 2014-03-20 final-report/paper.tex

--- 2014-03-20 final-report/paper.tex
@@
 
-\newcommand*{\system}{OGEN\xspace}
+\newcommand*{\system}{QuiViz\xspace}
 \newcommand*{\papertitle}{\system: Visualizing Physical Query Execution in a Relational Big Data Management System}

### 2014-03-20 final-report/paper.tex
Correct part about colors, remove word span

--- 2014-03-20 final-report/paper.tex
@@
 In particular, \system is built to inspect query execution in
-Myria~\cite{myria}, a distributed big data
+Myria\footnote{\url{http://myria.cs.washington.edu/}}, a distributed big data
 management system currently being developed in the UW CSE database group. Myria
@@
 saving them significant guesswork and programming effort to identify the appropriate logs to analyze. Our visualization tool,
-\system, uses profiling logs from Myria as the prototype DDMS. Myria is a distributed
+\system, uses profiling logs from \emph{Myria} \cite{myria} as the prototype DDMS. Myria is a distributed
 database platform that provides \emph{big data management and analytics as a service} primarily for scientific applications.
@@
 
-Ambrose\cite{ambrose} is a platform developed by Twitter to visualize and
-real-time monitor MapReduce data workflows. Ambrose
-offers three different views to show associated jobs, job dependencies and
-progress, which are not suitable for our needs as the abstraction level of jobs
-is too high and does not capture single operators.
+Twitter developed Ambrose\cite{ambrose}, a platform for visualization and real-time monitoring of MapReduce data workflows.
+They offer three different views to show associated jobs, job dependencies and progress.
+Ambrose has been released as open source.
+Ambrose, however, is not suitable for our needs as the abstraction level of jobs is too high and does not capture single operators.
 
-Google's Dapper\cite{sigelman2010dapper} is a distributed systems tracing
-infrastructure that offers fine-grained tracing of calls in Google's distributed
-systems. They also proposed an interface for visualizing traces. Similarly,
-X-trace\cite{fonseca2007x} was developed as a framework to trace which events
-cause what other events in a distributed environment. Recently, there has been
-work on visualizing event traces collected in
-X-trace\footnote{\url{https://github.com/brownsys/X-Trace/tree/master/src/webui/html/interactive}}.
-Due to the importance of these kinds of debug facilities, Twitter closely
-modeled Zipkin\cite{zipkin} after Dapper and X-trace and released it as open
-source.
+Google's Dapper\cite{sigelman2010dapper} is a distributed systems tracing infrastructure and offers fine grained tracing of calls in Google's distributed systems.
+They also proposed an interface for visualizing traces.
+Similarly, X-trace\cite{fonseca2007x} was developed as a framework to trace which events cause what other events in a distributed environment.
+recently, there has been work on visualizing event traces collected in X-trace\footnote{\url{https://github.com/brownsys/X-Trace/tree/master/src/webui/html/interactive}}.
+Due to the importance of these kinds of debug facilities, Twitter closely modeled Zipkin\cite{zipkin} after Dapper and X-trace and released it as open source.
 
-Dapper and X-trace focus on how data flows through a distributed system. In
-contrast, in \system the visualization focuses on the operators and how the
-data flows through them. This orthogonal view is better suited for debugging
-performance bottlenecks in DDBMSs and also scales to a larger number of events.
-Furthermore, \system offers different abstraction levels, which enables users
-to find problems faster and handle larger amounts of profiling data. Also,
-\system is specifically designed to help developers understand query execution
-in distributed database system as opposed to general traces in distributed
-systems.
+Dapper and X-trace focus on how data flows through a distributed system.
+In \system the visualization focuses on the operators and how the data flows through them.
+This orthogonal view is better suited for debugging performance bottlenecks in DDBMS and also scales to a larger number of events.
+Furthermore, in contrast to  we offer different abstraction levels, which enables users to find problems faster and handle larger amounts of profiling data.
+Also, \system is specifically designed to help developers understand query execution in distributed database system like Myria as opposed to general traces in distributed systems.
 
-Tools to visualize query plans, for example those used to improve the
-performance for the SDSS Sky survey\cite{szalay2002sdss}, focus on optimizing
-queries and not query execution and have no visualization of data flow, which
-is necessary to optimize physical query execution.
+Tools to visualize query plans for example in SQL Server, as used to improve performance for the SDSS Sky survey\cite{szalay2002sdss}, focus on optimizing queries and not query execution and have no visualization of data flow, which is necessary to optimize physical query execution.
 
@@
 
-\todo{use the word spans when referring to events with duration (used in dapper)}
-
 \subsubsection{Overview over all fragments}
@@
 The Worker Communication view consists of two charts: the aggregate communication
-chart and the time series chart. The aggregate communication chart is a matrix representing
+chart and the time series chart. The aggregate communication chart is a matrix (Figure~\ref{fig:network}) representing
 a cluster of $N$ workers. Each cell $m,n$ in the matrix represents the total number of tuples sent
-by worker $m$ to worker $n$ during the query execution. Figure~\ref{fig:matrix} shows this chart along
-with the selected link in the query plan DAG. Hovering the mouse over a cell brings up a tooltip
-stating the total number of tuples exchanged between the source and destination workers. The intensity
-of the cell color reflects the number of tuples exchanged, i.e., a larger number translates into a darker color
-and a smaller number results in a lighter color. The color pallet used is grayscale and hence
-immune to color-blindness issues. The color scale uses LAB interpolation.
-The matrix chart representing aggregate communication
-between workers help uncover data skew issues, as discussed in \S \ref{sec:eval}.
+by worker $m$ to worker $n$ during the query execution.
+We chose to use the colors from color brewer\cite{harrower2003colorbrewer} and interpolated the colors between the steps using the \emph{CIE L*a*b*} interpolation and lightness correction \cite{chromajsaisch}.
+The colors work in grayscale and can be interpreted by color blind users.
 
-We initially considered having a chord diagram\footnote{\url{http://bl.ocks.org/mbostock/4062006}} instead of a matrix
-for the aggregate communication chart. But feedback during the progress presentation session convinced us that
-a chord diagram would suffer from scalability problems. Adding an option for displaying a chord diagram remains
-part of future work.
+The matrix chart representing aggregate communication between workers help uncover data skew
+issues, as discussed in \S \ref{sec:eval}.
+
+% We initially considered having a chord diagram\footnote{\url{http://bl.ocks.org/mbostock/4062006}} instead of a matrix
+%for the aggregate communication chart. But feedback during the progress presentation session convinced us that
+%a chord diagram would suffer from scalability problems. Adding an option for displaying a chord diagram remains
+%part of future work.
 
@@
 
-% Adriana
-
 This paper introduced \system, a visualization tool carefully tailored to help
@@
 this issue. One example would be a more interactive query plan graph, which the
-users can zoom. Another example, in the \fragment view that shows spans
-for all events we should limit the data that is being downloaded and rendered
-to what the user can see.
+users can zoom. Also, using the ability to query the log data on the backend,
+the \fragment view should only download and render what the user can see.
 

### 2014-03-20 final-report/paper.tex
refactoring intro

--- 2014-03-20 final-report/paper.tex
@@
 
-We propose \system, a visualization tool that helps database developers/users
-explore and understand query execution and data movement in a distributed
-database management system (DDBMS). \system provides quick insight into common
-problems such as data skew or performance bottleneck by leveraging
-visualization techniques to present: (1) data flow between query segments and
-between workers, (2) query execution and operator dependencies, (3) cluster
-utilization, (4) network utilization.
+We propose a visualization system to explore and understand query
+execution and data movement in a distributed database management system
+(DDBMS). Our tool provides insight into: (1) data flow between query segments
+and between workers, (2) query execution and operator dependencies, (3) cluster
+utilization, (4) network utilization. Our tool, \system, helps
+developers understand and improve query execution, and provides insight into
+common problems such as data skew or performance bottlenecks.
 
@@
 
-Database query performance is often unpredictable, especially when the DBMS is
-distributed. This requires programmers or database administrators to spend
+Database query performance is often unpredictable, especially when the database is
+distributed. This requires programmers or database developers to spend
 significant amounts of time and effort isolating and debugging the root cause
@@
 
-Our goal is to build a visualization tool designed to help a database administrator understand the data flow and run-time
+Our goal is to build a visualization tool designed to help a database developer understand the data flow and run-time
 performance during query execution in a Distributed Database Management System (DDBMS). Using our tool, the
-database programmer or administrator could quickly discover run-time bottlenecks as well as problematic data skew issues,
+database programmer or developer could quickly discover run-time bottlenecks as well as problematic data skew issues,
 saving them significant guesswork and programming effort to identify the appropriate logs to analyze. Our visualization tool,
-\system, uses profiling logs from \emph{Myria} \cite{myria} as the prototype DDMS. Myria is a distributed
+\system, uses profiling logs from \emph{Myria} \cite{myria} as the prototype DDBMS. Myria is a distributed
 database platform that provides \emph{big data management and analytics as a service} primarily for scientific applications.
@@
 them in parallel. Currently, Myria does not offer a visualization tool designed to show the physical query plan, job execution and network
-communication between distributed workers. \system allows the programmer/database administrator to visualize the following aspects
-of query execution.
-\todo{it's not a dag, as we noticed}
+communication between distributed workers. \system is designed to help the Myria programmer answer the following questions regarding
+their query execution.
 
 \begin{itemize}
+\item \emph{How does the data flow through the system between operators?} 
+The overall data flow in a DDBMS is represented by a \graph. It is important
+for the programmer to understand the \graph in order to optimize it. To aid in this
+understanding, we provide an interactive visualization of the \graph in \system.
 
-\item An interactive visualization of the physical query plan. The physical
-query plan is composed of a DAG of query \emph{fragments}. A query
-fragment is a subtree of the query execution plan that is separated from the
-rest of the plan by a network boundary. This means that the operators inside a
-query fragment can be executed by a single thread. Each fragment is scheduled
-to run on multiple workers, which means each worker thread executes the same set of
-operators inside the fragment, but on different data. Each DAG link represents
-communication between workers, possibly over the network.  \system displays
-this DAG in an interactive fashion with clickable elements for fragment nodes
-and communication links.  This DAG also represents a high-level data flow in
-the DDBMS.
+%The physical
+%query plan is composed of a directed graph of query \emph{fragments}. A query
+%fragment is a subtree of the query execution plan that is separated from the
+%rest of the plan by a network boundary. This means that the operators inside a
+%query fragment can be executed by a single thread. Each fragment is scheduled
+%to run on multiple workers, which means each worker thread executes the same set of
+%operators inside the fragment, but on different data. Each graph link represents
+%communication between workers, possibly over the network.  \system displays
+%this \graph in an interactive fashion with clickable elements for fragment nodes
+%and communication links.  The \graph also represents a high-level data flow in
+%the DDBMS.
 
@@
 
-The role of the back-end is produce event logs that record when operators are called, when the call returns and when data is sent to another worker.
-These logs are then used to create the visualizations in \system's web UI as described in Section~\ref{sec:front}.
+The task of the back-end is to log when operators are called, when the call returns and when data is sent to another worker.
+Using this data and simple transformations, we can obtain the information necessary to create the visualizations to answer the questions in Section~\ref{sec:approach}.
 
-There are two possible approaches that we explored to collect event logs.
-The first and simplest approach was to write the logs on files directly to disk.
-The files would then get hauled back on one node using remote copy.
-The logs would then be parsed, and the data transformed into the desired format.
+There are two possible approaches that we explored to collect event logs that can be used to extract spans and other performance related data necessary for the visualization.
+First, events could be logged to files using a standard logging tool and then collect the files on one node using remote copy.
+Once the logs are on one machine, the logs can be parsed and the data can be transformed into the desired format.
 
-We switched to a more scalable data collection approach approach as shown in Figure~\ref{fig:arch} where logs get written directly into the worker's database as tuples in a relation.
-When the front-end client requests data, the master executes a query that collects the data and streams the results directly to the requesting client.
-We found this data collection approach to be more scalable, reliable and faster.
-More particularly, we are able to write queries to aggregate, filter and transform the collected data.
+We switched to a different approach as shown in Figure~\ref{fig:arch} where logs are written directly into the worker's database.
+When the front-end requests the data, the master executes a query that collects the data and streams the results directly to the requesting client.
+Logs are thus tuples in a relation and no longer separate in files and in a custom format.
+Despite being more reliable, faster and using the same logic that Myria already implements, we are able to write queries to aggregate, filter and transform the collected data.
 
-For example, when querying how many workers are working on a certain fragment at a given time to produce the visualization described in Section~\ref{sec:fragment}, we used a query to select only events in the root operators (the operators that have no parent in the same fragment) for each fragment and a custom map function that carries state.
+For example, the data for how many workers are working on a certain fragment as described in Section~\ref{sec:fragments} is generated from the event log using a query to select only events for the root operators (the operators that have no parent in the same fragment) in each fragment and a custom map function that carries state.
 The state is an integer that is incremented when the operator is called on a worker and decremented when the operator returns.
@@
 
+We highly suggest that developers use the second approach when porting \system to a new DDBMS.
+
 \subsection{Front-end}
-\label{sec:front}
 
@@
 D3 is a JavaScript framework for data visualization on the web.
+We used visualization techniques such as \emph{focus+context}\cite{furnas1986generalized} that possible to implement conveniently using D3.
 
 \system's front-end gets data from the Myria server as JSON and CSV files.
-JSON is used to describe the physical queryplan.
+JSON objects are used to describe the physical queryplan.
 CSV files contain logs on query execution at the fragment and operator level as well as data exchange between workers at different stages of the query execution.
@@
 
+\todo{use the word spans when referring to events with duration (used in dapper)}
+
 \subsubsection{Overview over all fragments}
-\label{sec:overview}
+\label{sec:fragments}
 
@@
 The Worker Communication view consists of two charts: the aggregate communication
-chart and the time series chart. The aggregate communication chart is a matrix (Figure~\ref{fig:network}) representing
+chart and the time series chart. The aggregate communication chart is a matrix representing
 a cluster of $N$ workers. Each cell $m,n$ in the matrix represents the total number of tuples sent
-by worker $m$ to worker $n$ during the query execution.
-We chose to use the colors from color brewer\cite{harrower2003colorbrewer} and interpolated the colors between the steps using the \emph{CIE L*a*b*} interpolation and lightness correction \cite{chromajsaisch}.
-The colors work in grayscale and can be interpreted by color blind users.
+by worker $m$ to worker $n$ during the query execution. Figure~\ref{fig:matrix} shows this chart along
+with the selected link in the query plan DAG. Hovering the mouse over a cell brings up a tooltip
+stating the total number of tuples exchanged between the source and destination workers. The intensity
+of the cell color reflects the number of tuples exchanged, i.e., a larger number translates into a darker color
+and a smaller number results in a lighter color. The color pallet used is grayscale and hence
+immune to color-blindness issues. The color scale uses LAB interpolation.
+The matrix chart representing aggregate communication
+between workers help uncover data skew issues, as discussed in \S \ref{sec:eval}.
 
-The matrix chart representing aggregate communication between workers help uncover data skew
-issues, as discussed in \S \ref{sec:eval}.
-
-% We initially considered having a chord diagram\footnote{\url{http://bl.ocks.org/mbostock/4062006}} instead of a matrix
-%for the aggregate communication chart. But feedback during the progress presentation session convinced us that
-%a chord diagram would suffer from scalability problems. Adding an option for displaying a chord diagram remains
-%part of future work.
+We initially considered having a chord diagram\footnote{\url{http://bl.ocks.org/mbostock/4062006}} instead of a matrix
+for the aggregate communication chart. But feedback during the progress presentation session convinced us that
+a chord diagram would suffer from scalability problems. Adding an option for displaying a chord diagram remains
+part of future work.
 
@@
 
+% Adriana
+
 This paper introduced \system, a visualization tool carefully tailored to help
@@
 this issue. One example would be a more interactive query plan graph, which the
-users can zoom. Also, using the ability to query the log data on the backend,
-the \fragment view should only download and render what the user can see.
+users can zoom. Another example, in the \fragment view that shows spans
+for all events we should limit the data that is being downloaded and rendered
+to what the user can see.
 

### 2014-03-20 final-report/paper.tex

--- 2014-03-20 final-report/paper.tex
@@
 
-Our goal is to build a visualization tool designed to help a database administrator understand the data flow and run-time
+Our goal is to build a visualization tool designed to help a database developer understand the data flow and run-time
 performance during query execution in a Distributed Database Management System (DDBMS). Using our tool, the
-database programmer or administrator could quickly discover run-time bottlenecks as well as problematic data skew issues,
+database programmer or developer could quickly discover run-time bottlenecks as well as problematic data skew issues,
 saving them significant guesswork and programming effort to identify the appropriate logs to analyze. Our visualization tool,
+\system, uses profiling logs from \emph{Myria} \cite{myria} as the prototype DDBMS. Myria is a distributed
+database platform that provides \emph{big data management and analytics as a service} primarily for scientific applications.
+Myria takes queries in high-level query languages, such as SQL, translating them into physical query plans and finally executing
+them in parallel. Currently, Myria does not offer a visualization tool designed to show the physical query plan, job execution and network
+communication between distributed workers. \system is designed to help the Myria programmer answer the following questions regarding
+their query execution. 
 \system, uses profiling logs from Myria as the prototype DDMS. Myria is a distributed
@@
 of query execution.
-\todo{it's not a dag, as we noticed}
 
 \begin{itemize}
+\item \emph{How does the data flow through the system between operators?} 
+The overall data flow in a DDBMS is represented by a \graph. It is important
+for the programmer to understand the \graph in order to optimize it. To aid in this
+understanding, we provide an interactive visualization of the \graph in \system.
 
-\item An interactive visualization of the physical query plan. The physical
-query plan is composed of a DAG of query \emph{fragments}. A query
-fragment is a subtree of the query execution plan that is separated from the
-rest of the plan by a network boundary. This means that the operators inside a
-query fragment can be executed by a single thread. Each fragment is scheduled
-to run on multiple workers, which means each worker thread executes the same set of
-operators inside the fragment, but on different data. Each DAG link represents
-communication between workers, possibly over the network.  \system displays
-this DAG in an interactive fashion with clickable elements for fragment nodes
-and communication links.  This DAG also represents a high-level data flow in
-the DDBMS.
+%The physical
+%query plan is composed of a directed graph of query \emph{fragments}. A query
+%fragment is a subtree of the query execution plan that is separated from the
+%rest of the plan by a network boundary. This means that the operators inside a
+%query fragment can be executed by a single thread. Each fragment is scheduled
+%to run on multiple workers, which means each worker thread executes the same set of
+%operators inside the fragment, but on different data. Each graph link represents
+%communication between workers, possibly over the network.  \system displays
+%this \graph in an interactive fashion with clickable elements for fragment nodes
+%and communication links.  The \graph also represents a high-level data flow in
+%the DDBMS.
 
@@
 correlations between the execution of different fragments. This view helps the user navigate between 
-fragments and is not intended to offer deep insight into performance issues. Section~{\sec:case1} describes
+fragments and is not intended to offer deep insight into performance issues. Section~\ref{sec:case1} describes
 a usecase for the \overall visualization.

### 2014-03-20 final-report/paper.tex
Clean up

--- 2014-03-20 final-report/paper.tex
@@
   \includegraphics[width=\columnwidth]{images/utilization_chart}
-  \caption{The utilization chart.}
+  \caption{The utilization chart. The user can brush on either the small graph on top or the bottom graph. Brushing on the second chart zooms the chart.}
   \label{fig:utilization_chart}
@@
 correlations between the execution of different fragments. This view helps the user navigate between
-fragments and is not intended to offer deep insight into performance issues. Section~{\sec:case1} describes
+fragments and is not intended to offer deep insight into performance issues. Section~\ref{sec:case1} describes
 a usecase for the \overall visualization.
@@
 The matrix chart representing aggregate communication between workers help uncover data skew
-issues, as discussed in \S \ref{sec:eval}.
-
-% We initially considered having a chord diagram\footnote{\url{http://bl.ocks.org/mbostock/4062006}} instead of a matrix
-%for the aggregate communication chart. But feedback during the progress presentation session convinced us that
-%a chord diagram would suffer from scalability problems. Adding an option for displaying a chord diagram remains
-%part of future work.
-
-\todo{is this figure necessary or is the figure in the evaluation section enough?}
-
-\begin{figure}[!ht]
-  \includegraphics[width=\columnwidth]{images/networkVis1}
-  \caption{Worker communication matrix chart. Each cell represents the number of exchanged tuples between a
-  pair of workers.}
-  \label{fig:matrix}
-\end{figure}
+issues, as discussed in Section~\ref{sec:eval}.
 

### 2014-03-20 final-report/paper.tex
programmer = developer

@@
 
-We propose \system, a visualization tool that helps database developers/users
+We propose \system, a visualization tool that helps database developers
 explore and understand query execution and data movement in a distributed
@@
 Database query performance is often unpredictable, especially when the database is
-distributed. This requires programmers or database developers to spend
+distributed. This requires database developers to spend
 significant amounts of time and effort isolating and debugging the root cause
@@
 performance during query execution in a Distributed Database Management System (DDBMS). Using our tool, the
-database programmer or developer could quickly discover run-time bottlenecks as well as problematic data skew issues,
+database developer could quickly discover run-time bottlenecks as well as problematic data skew issues,
 saving them significant guesswork and programming effort to identify the appropriate logs to analyze. Our visualization tool,
@@
 them in parallel. Currently, Myria does not offer a visualization tool designed to show the physical query plan, job execution and network
-communication between distributed workers. \system is designed to help the Myria programmer answer the following questions regarding
+communication between distributed workers. \system is designed to help the Myria developer answer the following questions regarding
 their query execution. 
@@
 them in parallel. Currently, Myria does not offer a visualization tool designed to show the physical query plan, job execution and network
-communication between distributed workers. \system allows the programmer/database administrator to visualize the following aspects
+communication between distributed workers. \system allows the user to visualize the following aspects
 of query execution.
@@
 The overall data flow in a DDBMS is represented by a \graph. It is important
-for the programmer to understand the \graph in order to optimize it. To aid in this
+for the developer to understand the \graph in order to optimize it. To aid in this
 understanding, we provide an interactive visualization of the \graph in \system.
@@
 several such causes, that we will enumerate below, we leveraged visualization
-techniques that we found fit to allow developers/programmers to get insight
+techniques that we found fit to allow developers to get insight
 into how the query was executed on the available resources and quickly
@@
 partitioning, workers might need to send unnecessary big amount of data between
-each other. We designed a view that allows the developer/programmer to analyze
+each other. We designed a view that allows the developer to analyze
 the data traffic generated while executing the query.

### 2014-03-20 final-report/paper.tex
Fix abstract

@@
 problems such as data skew or performance bottleneck by leveraging
-visualization techniques to present: (1) data flow between query segments and
+visualization techniques to present: (1) data flow between query operators and
 between workers, (2) query execution and operator dependencies, (3) cluster
@@
 
-In particular, \system is built to inspect query execution in
-Myria~\cite{myria}, a distributed big data
-management system currently being developed in the UW CSE database group. Myria
-aims towards building a distributed database platform to provide \emph{big data
-management and analytics as a service} primarily for scientific applications.
-The proposed visualizations can easily be applied to other DDBMS as well (e.g.
-Spark, Hadoop).
+In particular, \system is built to inspect query execution in Myria~\cite{myria},
+a distributed big data management system currently being developed in the
+UW CSE database group. Myria aims towards building a distributed database
+platform to provide \emph{big data management and analytics as a service} primarily
+for scientific applications. The proposed visualizations can easily be applied to
+other DDBMS as well (e.g. Spark, Hadoop).
 

### 2014-03-20 final-report/paper.tex
intro

--- 2014-03-20 final-report/paper.tex
@@
 
-\item A view for cluster utilization and operator performance for each query fragment. Each fragment is composed of
-multiple sequential database operators running on the same worker. Upon clicking on a collapsed fragment node in the
-query plan DAG, the node expands to display the operators comprising the fragment and renders the fragment execution view.
-The fragment execution view visualizes the utilization of the cluster by showing whether each worker is busy working
+\item \emph{How does the cluster visualization vary for each fragment and operator?} 
+%A view for cluster utilization and operator performance for each query fragment. 
+In \emph{Myria}, a fragment is composed of multiple sequential database operators running on the same worker. 
+A programmer will want to understand the cluster utilization at a given time as well as how much time is being spent
+executing an operator. \system provides a \fragment view to aid this understanding. 
+%Upon clicking on a collapsed fragment node in the
+%query plan DAG, the node expands to display the operators comprising the fragment and renders the fragment execution view.
+The \fragment view visualizes the utilization of the cluster by showing whether each worker is busy working
 on the selected fragment. It also breaks down the work done by each worker into operators composing the fragment.
@@
 represents source workers sending tuples to destination workers. Upon clicking on a link, the visualization renders the
-worker communication view. This shows the aggregate number of tuples exchanged between workers as well as a time series
+\network view. This shows the aggregate number of tuples exchanged between workers as well as a time series
 charting the number of tuples exchanged over time.
@@
 A query plan is an ordered set of steps used to access data in database management systems.
-The physical query plan is represented by a graph where each node represents a query fragment, and each link represents inter-worker communication from one query fragment to the next.
+The physical query plan is represented by a graph where each node represents a query fragment, and each link represents inter-\network from one query fragment to the next.
 Each query fragment is a collection of query operators.
@@
 When a fragment is selected, the operands inside of the fragment in the \graph view are color-coded to allow the user to easily match each task in the per-worker task schedule in the performance window with the corresponding query operator in the browser window.
-    \item \textbf{Edge selection:} upon selecting one or more fragment-to-fragment edges, a \network visualization gets displayed, providing information on inter-worker communication from one query fragment to the next.
+    \item \textbf{Edge selection:} upon selecting one or more fragment-to-fragment edges, a \network visualization gets displayed, providing information on inter-\network from one query fragment to the next.
 
@@
   \includegraphics[width=\columnwidth]{images/utilization_chart}
-  \caption{The utilization chart. The user can brush on either the small graph on top or the bottom graph. Brushing on the second chart zooms the chart.}
+  \caption{The utilization chart.}
   \label{fig:utilization_chart}
@@
 
-\subsubsection{Worker communication view}
+\subsubsection{\network view}
 
-The Worker Communication view consists of two charts: the aggregate communication
+The \network view consists of two charts: the aggregate communication
 chart and the time series chart. The aggregate communication chart is a matrix (Figure~\ref{fig:network}) representing
@@
 The matrix chart representing aggregate communication between workers help uncover data skew
-issues, as discussed in Section~\ref{sec:eval}.
+issues, as discussed in \S \ref{sec:eval}.
+
+% We initially considered having a chord diagram\footnote{\url{http://bl.ocks.org/mbostock/4062006}} instead of a matrix
+%for the aggregate communication chart. But feedback during the progress presentation session convinced us that
+%a chord diagram would suffer from scalability problems. Adding an option for displaying a chord diagram remains
+%part of future work.
+
+\todo{is this figure necessary or is the figure in the evaluation section enough?}
+
+\begin{figure}[!ht]
+  \includegraphics[width=\columnwidth]{images/networkVis1}
+  \caption{\network matrix chart. Each cell represents the number of exchanged tuples between a
+  pair of workers.}
+  \label{fig:matrix}
+\end{figure}
 
@@
   \includegraphics[width=\columnwidth]{images/networkVis2}
-  \caption{Worker communication time series chart. The user can select multiple cells in the matrix to compare the
+  \caption{\network time series chart. The user can select multiple cells in the matrix to compare the
   corresponding time series.}

### 2014-03-20 final-report/paper.tex

--- 2014-03-20 final-report/paper.tex
@@
 The matrix chart representing aggregate communication between workers help uncover data skew
-issues, as discussed in Section~\ref{sec:eval}.
+issues, as discussed in \S \ref{sec:eval}.
+
+We initially considered having a chord diagram\footnote{\url{http://bl.ocks.org/mbostock/4062006}} instead of a matrix
+for the \network chart. But feedback during the progress presentation session convinced us that
+a chord diagram would suffer from scalability problems. Adding an option for displaying a chord diagram remains
+part of future work.
+
+\todo{is this figure necessary or is the figure in the evaluation section enough?}
+
+\begin{figure}[!ht]
+  \includegraphics[width=\columnwidth]{images/networkVis1}
+  \caption{\network matrix chart. Each cell represents the number of exchanged tuples between a
+  pair of workers.}
+  \label{fig:matrix}
+\end{figure}
 

### 2014-03-20 final-report/paper.tex
intro changes

@@
 
-\item A view for communication between workers. A directed link between two fragments in the physical query plan DAG
-represents source workers sending tuples to destination workers. Upon clicking on a link, the visualization renders the
-\network view. This shows the aggregate number of tuples exchanged between workers as well as a time series
-charting the number of tuples exchanged over time.
+\item \emph{How is the input data partitioned?}
+Data skew is an important problem in a DDBMS, which can result in uneven communication between worker threads
+residing on different machines. \system provides a view for \network that represents communication   
+between two fragments in the \graph.
+
+%represents source workers sending tuples to destination workers. Upon clicking on a link, the visualization renders the
+%\network view. This shows the aggregate number of tuples exchanged between workers as well as a time series
+%charting the number of tuples exchanged over time.
 

### 2014-03-20 final-report/paper.tex
Sloppy

--- 2014-03-20 final-report/paper.tex
@@
 \begin{abstract}
-
-We propose \system, a visualization tool that helps database developers
-explore and understand query execution and data movement in a distributed
-database management system (DDBMS). \system provides quick insight into common
-problems such as data skew or performance bottleneck by leveraging
-visualization techniques to present: (1) data flow between query operators and
-between workers, (2) query execution and operator dependencies, (3) cluster
+\sloppy
+We propose \system, a visualization tool that helps database developers explore
+and understand query execution and data movement in a distributed database
+management system. \system provides quick insight into common problems
+such as data skew or performance bottleneck by leveraging visualization
+techniques to present: (1) data flow between query operators and between
+workers, (2) query execution and operator dependencies, (3) cluster
 utilization, (4) network utilization.
 
+\sloppy
 In particular, \system is built to inspect query execution in Myria~\cite{myria},

### 2014-03-21 final-report/paper.tex
m

@@
 
-\todo{is this figure necessary or is the figure in the evaluation section enough?}
+%\todo{is this figure necessary or is the figure in the evaluation section enough?}
 
@@
 time series chart below the matrix. The user can select multiple such cells to compare different time
-series. Figure~\ref{fig:ts} shows an example when the user selected two pairs: (8-2) and (7-6). The x-axis shows
+series. Figure~\ref{fig:ts} shows an example when the user selected two pairs: (8-1) and (7-5). The x-axis shows
 time in seconds and y-axis shows number of tuples exchanged between the source and destination workers.
 Just looking at the matrix chart it is evident that there is more data exchanged between the source-destination pair
-(8-2) than the pair (7-6). But as the time series comparison shows, there is more communication between (7-6)
-than (8-2) for the last 2.5 seconds of the query execution. This points to interesting patterns as to the rate at which
+(8-2) than the pair (7-6). But as the time series comparison shows, there is more communication between (7-5)
+than (8-1) for the last 2.5 seconds of the query execution. This points to interesting patterns as to the rate at which
 work is done at different workers at different stages of the execution. Distinct points on the lines are highlighted as

### 2014-03-21 final-report/paper.tex
new skew images
new execution skew image
new overview-skew figure
new utilization and operator charts
Final edits

--- 2014-03-20 final-report/paper.tex
@@
 \begin{abstract}
-\sloppy
-We propose \system, a visualization tool that helps database developers explore
-and understand query execution and data movement in a distributed database
-management system. \system provides quick insight into common problems
-such as data skew or performance bottleneck by leveraging visualization
-techniques to present: (1) data flow between query operators and between
-workers, (2) query execution and operator dependencies, (3) cluster
-utilization, (4) network utilization.
 
-\sloppy
+We propose \system, a visualization tool that helps database developers
+explore and understand query execution and data movement in a distributed
+database management system (DDBMS). \system provides quick insight into common
+problems such as data skew or performance bottleneck by leveraging
+visualization techniques to present: (1) data flow between query operators and
+between workers, (2) query execution and operator dependencies, (3) cluster
+utilization and (4) network utilization.
+
 In particular, \system is built to inspect query execution in Myria~\cite{myria},
@@
 platform to provide \emph{big data management and analytics as a service} primarily
-for scientific applications. The proposed visualizations can easily be applied to
+for scientific applications. \system was designed to be easily ported to
 other DDBMS as well (e.g. Spark, Hadoop).
@@
 Myria takes queries in high-level query languages, such as SQL, translating them into physical query plans and finally executing
-them in parallel. Currently, Myria does not offer a visualization tool designed to show the physical query plan, job execution and network
-communication between distributed workers. \system is designed to help the Myria developer answer the following questions regarding
-their query execution. 
-\system, uses profiling logs from Myria as the prototype DDMS. Myria is a distributed
-database platform that provides \emph{big data management and analytics as a service} primarily for scientific applications.
-Myria takes queries in high-level query languages, such as SQL, translating them into physical query plans and finally executing
-them in parallel. Currently, Myria does not offer a visualization tool designed to show the physical query plan, job execution and network
-communication between distributed workers. \system allows the user to visualize the following aspects
-of query execution.
+them in parallel.
+A query plan is an ordered set of steps used to access data in database management systems.
+Myria breaks down each query plan into \emph{fragments}.
+A fragment is composed of multiple sequential database operators running on one worker.
+Fragments map down to a unit of work that can be scheduled and parallelized across a cluster.
+
+Currently, Myria does not offer a detailed visualization front-end designed to provide insight into job execution and network
+communication between distributed workers. \system is designed to help the Myria developers write better optimized queries by addressing
+the following questions.
 
 \begin{itemize}
-\item \emph{How does the data flow through the system between operators?} 
-The overall data flow in a DDBMS is represented by a \graph. It is important
-for the developer to understand the \graph in order to optimize it. To aid in this
-understanding, we provide an interactive visualization of the \graph in \system.
+\item \emph{How does the data flow through the workers between query fragments?} 
+\item \emph{How can we identify data skew and execution skew?} 
+\item \emph{What causes performance bottlenecks?}
+\end{itemize}
 
-%The physical
-%query plan is composed of a directed graph of query \emph{fragments}. A query
-%fragment is a subtree of the query execution plan that is separated from the
-%rest of the plan by a network boundary. This means that the operators inside a
-%query fragment can be executed by a single thread. Each fragment is scheduled
-%to run on multiple workers, which means each worker thread executes the same set of
-%operators inside the fragment, but on different data. Each graph link represents
-%communication between workers, possibly over the network.  \system displays
-%this \graph in an interactive fashion with clickable elements for fragment nodes
-%and communication links.  The \graph also represents a high-level data flow in
-%the DDBMS.
-
-\item \emph{How does the cluster visualization vary for each fragment and operator?} 
-%A view for cluster utilization and operator performance for each query fragment. 
-In \emph{Myria}, a fragment is composed of multiple sequential database operators running on the same worker. 
-A programmer will want to understand the cluster utilization at a given time as well as how much time is being spent
-executing an operator. \system provides a \fragment view to aid this understanding. 
-%Upon clicking on a collapsed fragment node in the
-%query plan DAG, the node expands to display the operators comprising the fragment and renders the fragment execution view.
-The \fragment view visualizes the utilization of the cluster by showing whether each worker is busy working
-on the selected fragment. It also breaks down the work done by each worker into operators composing the fragment.
-
-\item A view for communication between workers. A directed link between two fragments in the physical query plan DAG
-represents source workers sending tuples to destination workers. Upon clicking on a link, the visualization renders the
-\network view. This shows the aggregate number of tuples exchanged between workers as well as a time series
-charting the number of tuples exchanged over time.
-
-\end{itemize}
+We show in the next sections how \system addresses these questions by the means of effective data visualization.
 
@@
 \section{Related Work}
+
+Performance visualization tools for distributed systems are not a new idea.
 
@@
 The first step in designing \system was to identify the possible causes that 
-could affect the execution performance of the query. After we identified
-several such causes, that we will enumerate below, we leveraged visualization
-techniques that we found fit to allow developers to get insight
-into how the query was executed on the available resources and quickly
-identify the source(s) of the performance penalty.
+affect the performance of a query execution. After we identified
+several such causes, we explored visualization
+techniques we found adequate in exposing those performance bottlenecks.
 
-We found that the performance of the query might be affected by the following
+We found that the performance of a query might be affected by the following
 factors, which greatly influenced the design of our visualization tool:
@@
    \item \emph{Stragglers, tail latency, expensive query operators, poor
-scheduling.} Our targeted systems are \emph{distributed} databases. Therefore,
-as the query is executed in a distributed environment, the query coordinator
+scheduling.} We target \emph{distributed} database management systems. Therefore,
+as the query gets executed in a distributed environment, the query coordinator
 has to wait until all the workers finish executing their assigned tasks. To
-give insight into how the tasks are distributed and executed at each workers,
-we designed a view that shows the cluster utilization over time and at each
-worker.
+give insight into how the tasks are distributed and executed on each worker,
+we provide a view that exposes utilization details at the cluster and worker levels.
    \item \emph{Poor data partitioning, data skew.} Network traffic can cause
 serious delay when executing a distributed query. Due to poor data
-partitioning, workers might need to send unnecessary big amount of data between
-each other. We designed a view that allows the developer to analyze
+partitioning, workers might need to send large amounts of data to one
+another. We designed a view that allows the developer to analyze
 the data traffic generated while executing the query.
@@
 % explain how events and logs are used to create the raw data for vis
+\subsection{System Overview}
 
 \system's architecture (Figure~\ref{fig:arch}) is composed of: (1) a back-end
-used as a plug-in interface for the targeted distributed database system, and (2) a
+used as a plug-in interface to the targeted distributed database system, and (2) a platform-independent
 front-end that produces performance visualizations what we describe in Section~\ref{sec:front}. The front-end and
@@
 
-The role of the back-end is produce event logs that record when operators are called, when the call returns and when data is sent to another worker.
+The role of \system's back-end is to produce event logs that include: when operators are called, when the call returns and when data is sent between workers.
 These logs are then used to create the visualizations in \system's web UI as described in Section~\ref{sec:front}.
 
-There are two possible approaches that we explored to collect event logs.
-The first and simplest approach was to write the logs on files directly to disk.
-The files would then get hauled back on one node using remote copy.
-The logs would then be parsed, and the data transformed into the desired format.
+We explored two approaches for collecting event logs.
+The simplest approach was to write the logs on files directly to disk.
+The files would then get hauled back to one master node using remote copy where they would then get parsed, manipulated to the desired format.
 
-We switched to a more scalable data collection approach approach as shown in Figure~\ref{fig:arch} where logs get written directly into the worker's database as tuples in a relation.
-When the front-end client requests data, the master executes a query that collects the data and streams the results directly to the requesting client.
-We found this data collection approach to be more scalable, reliable and faster.
-More particularly, we are able to write queries to aggregate, filter and transform the collected data.
+We switched to a more scalable data collection approach described in Figure~\ref{fig:arch} where logs get written directly into the local worker's database as tuples in a relation.
+When the front-end client requests data, the master executes a query that gathers the log data from the dirstributed workers, and streams the results directly to the requesting client.
+We found this data collection approach to be more scalable, reliable and faster than the simpler file-based logging approach.
+More particularly, using a database to manage performance logs allowed us to write queries to aggregate, filter and transform the collected data.
 
-For example, when querying how many workers are working on a certain fragment at a given time to produce the visualization described in Section~\ref{sec:fragment}, we used a query to select only events in the root operators (the operators that have no parent in the same fragment) for each fragment and a custom map function that carries state.
+For example, to produce the visualization described in Section~\ref{sec:fragment}, we had to query how many workers are working on a certain fragment at a given time.
+To do this, we used a query to select events in the root operators (i.e.
+parentless operator) of each fragment and used a custom map function to carry state.
 The state is an integer that is incremented when the operator is called on a worker and decremented when the operator returns.
@@
 D3 is a JavaScript framework for data visualization on the web.
-
-\system's front-end gets data from the Myria server as JSON and CSV files.
-JSON is used to describe the physical queryplan.
-CSV files contain logs on query execution at the fragment and operator level as well as data exchange between workers at different stages of the query execution.
 
@@
 
-The \graph view as pictured in Figure~\ref{fig:graph} allows the user to selectively browse different performance visualizations tied to a given query execution.
-The \graph view provides as the name indicates, a visualization of the physical query plan.
-A query plan is an ordered set of steps used to access data in database management systems.
-The physical query plan is represented by a graph where each node represents a query fragment, and each link represents inter-\network from one query fragment to the next.
-Each query fragment is a collection of query operators.
-One can think of a query fragment as a job unit that gets scheduled on a cluster of machines.
-Consequently transitioning from one query fragment to another generally leads to all-to-all communication between worker nodes on the cluster.
+The \graph view as pictured in Figure~\ref{fig:graph} allows the user to selectively browse different performance visualizations of a given query.
+The physical query plan is represented by a graph where each node represents a query fragment, and each link represents inter-worker communication between the execution of two query fragments.
+Each query fragment is composed by a collection of query operators that can be executed together as one job.
 
-The user can perform three classes of actions that will render a new visualization in the performance panel:
+The user can perform three classes of actions on the \graph view that will render a new visualization in the performance panel:
 \begin{itemize}
@@
 When a fragment is selected, the operands inside of the fragment in the \graph view are color-coded to allow the user to easily match each task in the per-worker task schedule in the performance window with the corresponding query operator in the browser window.
-    \item \textbf{Edge selection:} upon selecting one or more fragment-to-fragment edges, a \network visualization gets displayed, providing information on inter-\network from one query fragment to the next.
+    \item \textbf{Edge selection:} upon selecting one or more fragment-to-fragment edges, a \network visualization gets displayed, providing information on inter-worker communication transitioning from one query fragment to the next.
 
@@
 \textbf{Graph Rendering:} We use D3 \cite{2011-d3} to render the \graph graph, which allows us to support various interactions and transitions.
-We use GraphViz \cite{Ellson01graphviz} in the backend to generate graph layout information.
-GraphViz allows us to generate graph layouts that are optimized to minimize area footprint.
+We use GraphViz \cite{Ellson01graphviz} in the backend to generate graph layout data.
+GraphViz generates graph layouts that are optimized to minimize area footprint.
 The layout information is then fed into a D3-based rendering engine which supports various interactions and animations techniques.
@@
 The first mechanism can reduce the size of a graph by a constant factor by collapsing operators that compose a fragment into a single fragment node.
-The user can click once on a collapsed fragment to expand, and click once on an expanded fragment to collapse it.
+The user can click once on a collapsed fragment to expand it, and click once on an expanded fragment to collapse it.
 Expanding or reducing a node can cause large changes in the graph layout as GraphViz changes the graph layout to minimize overall area.
-This graph re-shuffling is illustrated in Figure~\ref{graph} where expanding Frag2 causes the layout to change.
+This graph re-shuffling is illustrated in Figure~\ref{graph} where expanding Fragment~2 causes the layout to change.
 To address these layout changes, we implemented transitions to allow the user to track the fragments as those get reshuffled.
@@
 The \fragment view comprises two charts: the utilization chart and the
-operators chart. The utilization chart (Figure~\ref{fig:utilization_chart}) shows on
-how many workers a fragment is running over time. The chart is used to
+operators chart. The utilization chart (Figure~\ref{fig:utilization_chart}) shows
+how many workers a fragment is running on over time. The chart is used to
 quickly reveal any patterns in the schedule, like tail latency, long periods of
-idleness, etc. Execution skew is visible as humps.
-We implemented \emph{focus+context}\cite{furnas1986generalized} in form of
+idleness, etc. 
+We implemented \emph{focus+context}\cite{furnas1986generalized} using 
 a small brush (at the top of the figure) that allows the user to zoom-in and further
-analyze the problematic area without loosing the context of the whole execution.
+analyze a problematic area without loosing context on the entire execution.
 
@@
   \includegraphics[width=\columnwidth]{images/utilization_chart}
-  \caption{The utilization chart. The user can brush on either the small graph on top or the bottom graph. Brushing on the second chart zooms the chart.}
+  \caption{The utilization chart. The user can brush on either the small graph on top or the bottom graph to zoom into a region of interest.}
   \label{fig:utilization_chart}
@@
 The utilization chart also implements a brush that allows the user to further
-analyze the query execution by displaying the schedule of each of the worker
-nodes, in the operators chart (Figure~\ref{fig:operators_chart}). The per-worker
-schedule shows what operators the worker executed and for how long.
+analyze query execution by displaying the work schedule of each worker 
+in the operators chart (Figure~\ref{fig:operators_chart}). The per-worker
+schedule displays what operators each worker is executing over time.
 
@@
 We chose to use the colors from color brewer\cite{harrower2003colorbrewer} and interpolated the colors between the steps using the \emph{CIE L*a*b*} interpolation and lightness correction \cite{chromajsaisch}.
-The colors work in grayscale and can be interpreted by color blind users.
+The colors work in grayscale and can be interpreted by color-blind users.
 
 The matrix chart representing aggregate communication between workers help uncover data skew
-issues, as discussed in \S \ref{sec:eval}.
+issues, as discussed in Section~\ref{sec:eval}.
 
 We initially considered having a chord diagram\footnote{\url{http://bl.ocks.org/mbostock/4062006}} instead of a matrix
-for the \network chart. But feedback during the progress presentation session convinced us that
+for the \network chart. But feedback from our peers convinced us that
 a chord diagram would suffer from scalability problems. Adding an option for displaying a chord diagram remains
@@
 
-When the user clicks on a matrix cell, they are able to view the breakdown of the worker pair communication in a
+When the user clicks on a matrix cell, she is able to view the breakdown of the worker pair communication in a
 time series chart below the matrix. The user can select multiple such cells to compare different time
@@
 than (8-2) for the last 2.5 seconds of the query execution. This points to interesting patterns as to the rate at which
-work is done at different workers at different stages of the execution. Distinct points on the lines are highlighted as
+work is executed on different workers during different stages of the execution. Distinct points on the lines are highlighted as
 black dots. Hovering the cursor over a dot shows the information for that data point in a tooltip. Meanwhile the cells
-in the matrix remain highlighted with the same color used to draw the corresponding time series line. This aids in the
-user keeping track of the mapping between the matrix cell and the corresponding time series line. We also provide
+in the matrix remain highlighted using the same color used to draw the corresponding time series line. This helps the
+user keep track of the mapping between the matrix cell and the corresponding time series line. We also provide
 a 'clear selection' button that erases everything from the time series chart upon clicking. Finally we provide the
@@
 
-We explore a simple join query we performed on one million tuples extracted from the Twitter user graph.
-This query joins Twitter followers and followees by user ID.
-The query can be written in SQL as:
+We explore a simple join query we performed on one million tuples sampled from a Twitter user graph. This query joins Twitter followers and followees by user ID. The query can be written in SQL as:
 
@@
 The resulting \graph visualization is displayed in Figure~\ref{fig:join}.
-The user will navigate the \graph visualization in the browser panel in the next case studies to identify several performance symptoms.
+The user will navigate the \graph visualization in the browser panel in the next case studies to identify different performance issues.
 
@@
 that Fragment~2 is irregularly executed on multiple workers. 
-If the graph is not monolithic (as it is the case in Fragment 1, 3 and 4), workers are sleeping because they
-either (1) don't have enough data available to work on (data dependency problem) or (2) the destination worker's
-input buffers are full thus forcing the source workers to stall.
+If the graph is not one solid block as it in Fragment 1, 3 and 4, one can assume workers are sleeping. Having workers sleep 
+can be caused by one of two reasons: (1) the workers don't have enough data available to work on (i.e.
+data dependency problem) or (2) the destination worker's input buffers are full thus forcing the source workers to stall (i.e.
+back-pressure problem).
 
@@
 
-\todo{Let's not call the fragments frag2 but write fragment 2}
-
 The \network visualization at a glance shows how much data was sent between workers by one query fragment to the next.
-The two \network views in Figure~\ref{fig:network} were obtained by selecting the Frag0$\rightarrow$Frag2 and Frag1$\rightarrow$Frag2 edges respectively in the \graph view.
+The two \network views in Figure~\ref{fig:network} were obtained by selecting the Fragment0$\rightarrow$Fragment2 and Fragment1$\rightarrow$Fragment2 edges respectively in the \graph view.
 Figure~\ref{fig:network} displays aggregate data communication between workers.
-In Figure~\ref{fig:no-skew} the amount of data sent from each worker to every other worker is balanced as opposed to Figure~\ref{fig:skew} where one can see that the volume of data sent to worker~1 and worker~5 outweighs data sent to the other workers in the system.
-This apparent data skew could be caused by partitioning issues.
+In Figure~\ref{fig:no-skew}, the amount of data sent from each worker to every other worker is balanced as opposed to Figure~\ref{fig:skew} where one can see that the volume of data sent to worker~1 and worker~5 dwarfs data sent to the other workers in the system.
+This apparent data skew could be caused by poor partitioning.
 
@@
 The \fragment visualization allows the user to dive down into query execution details at the worker level.
-The \fragment view in Figure~\ref{fig:exec_skew} was obtained by selecting Frag2, which is the fragment that performs the join, in the \graph view.
+The \fragment view in Figure~\ref{fig:exec_skew} was obtained by selecting Fragment2 in the \graph view.
+Fragment2 is the fragment that performs the join.
 The user can observe at a glance that certain workers produce much more data as a result of the join execution.
-Particularly, we saw in Case 2 Figure~\ref{fig:skew}, that tuples sent to Fragment 2 from Fragment 1 were heavily skewed towards being sent to worker~1 and worker~5.
+Particularly, we saw in Case 2 Figure~\ref{fig:skew}, that tuples sent to Fragment2 from Fragment1 were heavily skewed towards being sent to worker~1 and worker~5.
 In Figure~\ref{fig:exec_skew}, the user can now see that as a result of the join, worker~5 ends up producing most tuples that have to be written back to disk.
@@
 In the short term, the user can modify the way the input data set gets partitioned.
-Alternatively she could change the number of workers allocated for the job.
-in the long term, developers should look into ways to improve data flow to avoid that workers have to wait for input data to process.
+Alternatively the user could change the number of workers allocated for the job.
+In the long term, developers should look at ways to improve data flow to minimize idling on workers that are waiting on input data.
 
@@
 
-All in all, \system allows the user to get much more insight into the execution of a query than what was available before, i.e.
-text-based performance logs, and runtimes.
-Being able to navigate performance data using a simple interface can indeed facilitate the task of a developer who is designing Myria's back end, or a Myria user who is writing complex queries to process large amounts of data.
-That said, \system can only help the user visualize query execution performance, but won't necessarily point the user to what she needs to do to improve performance of her query execution.
+All in all, \system allows the user to get much more insight into query execution than what was available before, i.e.
+text-based performance logs, and runtimes data.
+Being able to navigate performance data using a simple visual interface can indeed facilitate the task of a developer who is designing Myria's back end, or a Myria user who is writing complex queries to process large amounts of data.
+That said, \system can only help the user visualize query execution performance.
+\system but won't necessarily point the user to what needs to be done to improve query performance.
+We believe that an experienced user will quickly learn how to identify symptomatic performance execution, and know what steps to take to improve performance as a result.
+We would have to back this up with a user-study, which could be the object of future work in this direction.
 
@@
 This paper introduced \system, a visualization tool carefully tailored to help
-database developers or users to quickly identify common distributed databases
-performance bottlenecks, like tail latencies, data skew, etc. Our tool has been
-successfully integrated with the Myria distributed database system, and has already
-been used to identify several bugs, like poor physical query plan optimization,
-tail latencies, poor data partitioning, which led us to believe the
-visualization techniques we used are appropriate to help the user quickly
-navigate to/find the cause of the performance problem. \system could be easily
-plugged-in with another distributed database system, the only requirement being
+database developers and users to quickly identify common distributed databases
+performance bottlenecks, such as stragglers, data skew, etc. Our tool has been
+successfully integrated with the Myria distributed database management system, and has already
+been used to identify several bugs, such as  poor physical query plan optimization,
+stragglers, poor data partitioning etc. This has led us to believe that the
+visualization techniques we employed are effective in helping the user quickly
+narrow down and identify the cause of the performance bottleneck. \system could be easily
+integrated with another distributed database system, the only requirement being
 a specific log data format. We explained how we collect logs in Myria and recommend
@@
 The current implementation focuses on scalability. We managed to some extent to
-address the scalability issues raised by big query plans, large amount of
+address some scalability issues raised by big query plans, large amount of
 workers or long running queries. For example the \network view uses a
@@
 this issue. One example would be a more interactive query plan graph, which the
-users can zoom. Also, using the ability to query the log data on the backend,
+users can zoom into. Also, by querying the performance log data on the backend,
 the \fragment view should only download and render what the user can see.
@@
 visualizations described in this paper, it will form a powerful debugging tool
-that will help uers better understand and improve query execution.
+that will help users better understand and improve query execution.
 

### 2014-03-21 final-report/paper.tex

@@
 
-We propose \system, a visualization tool that helps database developers
-explore and understand query execution and data movement in a distributed
-database management system (DDBMS). \system provides quick insight into common
-problems such as data skew or performance bottleneck by leveraging
-visualization techniques to present: (1) data flow between query operators and
-between workers, (2) query execution and operator dependencies, (3) cluster
-utilization and (4) network utilization.
+\sloppy
+We propose \system, a visualization tool that helps database developers explore
+and understand query execution and data movement in a distributed database
+management system. \system provides quick insight into common problems
+such as data skew or performance bottleneck by leveraging visualization
+techniques to present: (1) data flow between query operators and between
+workers, (2) query execution and operator dependencies, (3) cluster
+utilization, (4) network utilization.
 
+\sloppy
 In particular, \system is built to inspect query execution in Myria~\cite{myria},
@@
 \begin{itemize}
-\item \emph{How does the data flow through the workers between query fragments?} 
-\item \emph{How can we identify data skew and execution skew?} 
-\item \emph{What causes performance bottlenecks?}
+\item \emph{How does the data flow through the system between operators?} 
+The overall data flow in a DDBMS is represented by a \graph. It is important
+for the developer to understand the \graph in order to optimize it. To aid in this
+understanding, we provide an interactive visualization of the \graph in \system.
+
+\item \emph{How does the cluster visualization vary for each fragment and operator?} 
+In \emph{Myria}, a fragment is composed of multiple sequential database operators running on the same worker. 
+A programmer will want to understand the cluster utilization at a given time as well as how much time is being spent
+executing an operator.
+\system provides a \fragment view to aid this understanding.
+The \fragment view visualizes the utilization of the cluster by showing what operator in the fragment each worker is busy working on.
+
+\item \emph{How is the input data partitioned?}
+Data skew is an important problem in a DDBMS, which can result in uneven communication between worker threads
+residing on different machines. \system provides a view for \network that all-to-all communication   
+between two fragments in the \graph.
+
 \end{itemize}
@@
 \textbf{Graph Navigation:} A query plan can be arbitrarily large.
-Thus we offer two mechanisms that facilitate exploration of the graph for the user: (1) expanding/reducing fragments and (2) zooming.
+Thus we offer two mechanisms that facilitate exploration of the graph for the user: (1) expanding/reducing fragments and (2) paning.
 The first mechanism can reduce the size of a graph by a constant factor by collapsing operators that compose a fragment into a single fragment node.
@@
 To address these layout changes, we implemented transitions to allow the user to track the fragments as those get reshuffled.
-The second mechanism that facilitates exploration of the graph for the user is zooming.
-This feature was implemented in D3 and allows the user to quickly jump from one part of the graph to the next, by zooming in and out of the graph view.
+The second mechanism that facilitates exploration of the graph for the user is paning.
+This feature was implemented in D3 and allows the user to navigate a graph when it doesn't fit inside of the browser panel.
 
@@
 \begin{figure}[ht]
-  \includegraphics[width=\columnwidth]{images/operators_chart}
+  \includegraphics[width=\columnwidth]{images/operators-chart}
   \caption{The operators chart.}
@@
 time series chart below the matrix. The user can select multiple such cells to compare different time
-series. Figure~\ref{fig:ts} shows an example when the user selected two pairs: (8-2) and (7-6). The x-axis shows
+series. Figure~\ref{fig:ts} shows an example when the user selected two pairs: (8-1) and (7-5). The x-axis shows
 time in seconds and y-axis shows number of tuples exchanged between the source and destination workers.
 Just looking at the matrix chart it is evident that there is more data exchanged between the source-destination pair
-(8-2) than the pair (7-6). But as the time series comparison shows, there is more communication between (7-6)
-than (8-2) for the last 2.5 seconds of the query execution. This points to interesting patterns as to the rate at which
-work is executed on different workers during different stages of the execution. Distinct points on the lines are highlighted as
+(8-2) than the pair (7-6). But as the time series comparison shows, there is more communication between (7-5)
+than (8-1) for the last 2.5 seconds of the query execution. This points to interesting patterns as to the rate at which
+work is done at different workers at different stages of the execution. Distinct points on the lines are highlighted as
 black dots. Hovering the cursor over a dot shows the information for that data point in a tooltip. Meanwhile the cells
@@
 \begin{figure}[ht]
-  \includegraphics[width=\columnwidth]{images/overview_skew}
+  \includegraphics[width=\columnwidth]{images/overview-skew}
   \caption{This \overall shows very irregular usage in fragment 2.}
@@
   \begin{subfigure}[b]{0.49\columnwidth}
-    \includegraphics[width=\columnwidth]{images/skew}
+    \includegraphics[width=\columnwidth]{images/skew.pdf}
     \caption{Data skew.}
@@
   \begin{subfigure}[b]{0.49\columnwidth}
-    \includegraphics[width=\columnwidth]{images/no-skew}
+    \includegraphics[width=\columnwidth]{images/no-skew.pdf}
     \caption{No data skew.}
@@
 \begin{figure}[ht]
-  \includegraphics[width=\columnwidth]{images/execution_skew}
+  \includegraphics[width=\columnwidth]{images/execution-skew}
   \caption{This \fragment view shows heavy execution skew on worker~5.}
